---
title: "Individual homework 5"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, error = TRUE)
```

## Please push your assignment to GitHub by 10:00 pm (Central) Tuesday, May 7. 


You are currently in the GitHub repository (repo) for `hw5-username`. The assignment prompt is shown below (i.e. in `README.Rmd`). You can view this online in your homework 5 GitHub repository as a Markdown file(`README.md`) or a pdf.

Please **use `hw5.Rmd` to complete this assignment**. Be sure to **knit your file to PDF before your final push to GitHub**.


## Homework process

For help on the homework process, review

-   [Assignments in Stat 220](https://stat220-spring24.netlify.app/assignments) for
    content/formatting questions.

-   [GitHub Guide for Students in Stat 220](https://stat220-spring24.netlify.app/github_tutorial)
    for Git and Github instructions.

When you are done with your homework, **don’t forget to push your changes to ALL files back to GitHub!** This means you should commit and push all related files, not just your final PDF. Additionally, ensure you post the link to your **GitHub repository to Gradescope for the final submission and grading**. This step is crucial as it allows for a comprehensive review of both your code and the rendered output.

------------------------------------------------



## Problem 1: regular expression using `babynames`

### a. 

Construct a regular expression (regex) to find all names that end in a vowel (here, you can consider "y" to be a vowel). Store this regex in `pattern` and run the code to determine how many baby names in 2017 ended with a vowel. Use babynames dataset from the `babynames` package.



### b. 

Write a regex that finds names that start in "Ed" and ends with "rd" so that it matches names like `"Edward"` and `"Eddard"`. Any number of characters can be in between. Check your regex using the small vector given below then determine how many babynames in 2017 match this pattern.

```{r}
x <- c("Edward", "Eddard", "Ned")
```



### c.

Construct a regex to find all names that only contain vowels in 2017. (In this problem, consider vowels to be aeiouy.) Are there any such names? If so, what are they?


### d.

Construct a regex to find all four-letter names in 2017 and store these names in a vector. (Hint: try using `dplyr::pull`) Print the length of this vector. (Please don't print the entire vector.)



### e.

Write a regex to find all of the four letter names that are palindromes in your names from part (d). ("Anna" is one such example.) Check this on the baby names from 2017. (Hint: You will need to capture the first two letters individually and then use back references. Also, it's easier to operate on all lowercase letters.)



----------------------------------------------------------



## Problem 2: Energy autocorrelation

Auto correlation measures the correlation between measurements that differ by a fixed amount of time. For example, "lag 1" autocorrelation measures the correlation between measurements that are 1 time unit apart, lag 2 measures the correlation between measurements 2 time units apart, etc. 

```{r}
energy <- readr::read_csv("https://raw.githubusercontent.com/deepbas/statdatasets/main/energy.csv",
                    col_type = cols(
                     .default = col_double(), 
                      Timestamp = col_datetime(format = ""),
                      dayWeek = col_factor(levels=c("Mon","Tues","Wed","Thurs","Fri","Sat","Sun"))
                     ))
```

The `acf` function computes autocorrelation in R. Here we get autocorrelation for Olin energy readings. The correlation between units 0 minutes apart is 1 (they are the same measurements!), the correlation between readings 15 minutes apart is 0.956, between 30 minutes apart is 0.95, between 45 minutes apart is 0.934 and between 60 minutes apart is 0.917. It makes sense that a correlation exists between energy use at points close in time. 

```{r}
x <- energy %>% 
  arrange(Timestamp) %>%    # making sure sorted by time
  pull("Olin_Hall_of_Science")
acf_out <- acf(
  x,   # time series
  na.action = na.pass,    # skips over NAs
  lag.max = 4,   # max lag
  plot = FALSE)   # don't plot
acf_out
acf_out$acf   # autocorr values
acf_out$lag   # lag values
```


### a.  

Write a function called `autocor_fun` that will take in:

- the vector of KWH values from one building
- a max lag value
  
and return a *data frame* with variables

- `autocor` which measures autocorrelation between `energyKWH` values. Note that you will need to extract the `acf` values from the `acf` function and coerce these into a vector with `as.vector`. 
- `lag` which tells what lag the `autocor` was computed at

Use the `na.action` and `plot` arguments in `acf` that are shown above when writing your function.

Check your `autocor_fun` on the Olin KWH data that was used above with a max lag of 4. 


### b. 

Use the `purrr` package to apply `autocor_fun` to the buildings `"Sayles-Hill" ,"Language_&_Dining_Center", "Olin_Hall_of_Science"` with a max lag of 4. Use the wide data frame `energy` and your result should be a data frame that also identifies buildings. 


### c. 

Repeat (b) but this time use the narrow version of the data `energy_narrow`. This time you will use `dplyr` groupings and summarize to apply the  `autocor_fun` to each building's KWH. (Don't forget to filter to the three buildings). Your data frame results for (b) and (c) should be identical.

```{r}
energy_narrow <- energy %>% 
  mutate(month = month(month, label=TRUE)) %>%
  pivot_longer(
    cols = 9:90,
    names_to = "building", 
    values_to = "energyKWH") 
```


### d. 

Create a data frame that contains acf values for 24 hours (i.e. a max lag of 24x4=96) for the three buildings used in (b, c). The plog acf values (y) against lag (x) for each of the buildings and describe what trends you see. 


--------------------------------------------


## Problem 3: weather

Load the `nasaweather` data and look at the help file for the atmospheric data `?atmos`. 

### a.
Create a `for` loop that computes the coefficient of variation for all atmospheric measurements except location (lat/long) and time (year, month) variables. The coefficient of variation is the ratio of the standard deviation over the mean of a variable. Print out the output of your loop.


### b.

Use a `map` function (from `purrr` package) to compute the coefficient of variation for all atmospheric measurements except location (lat/long) and time (year, month) variables. Use a function that returns a vector or data frame (and show this output).


### c. 

Create a function called `my_stats` that computes the following statistics for an input vector: mean, sd, and 5-number summary (min/Q1/median/Q3/max). The function should return an  output object that is a "named" vector. E.g. the following is a named vector with the name of the left side of `=` and the value on the right side. Show the output of your function by inputting the `temp` values from `atmos`.
```{r}
# e.g. named vector with names x and y and values 1 and 2
c(x = 1, y = 2)
```



### d. 

Use `map_df` to compute `my_stats` for all atmospheric measurements except location (lat/long) and time (year, month) variables. Include the variable names as a column in your data frame. 


### e. 

For each year, compute `my_stats` for the `temp` variable. To do this:

- use a year grouping and the `summarize` command 
- add a `stats` variable that identifies which statistic is entered in a given row
- then make a wider version of this data that contains a column for each of the `my_stats` for each year in the data.  (Final dimensions should be 6x8)


--------------------------------------------------------------


## Problem 4: String Extraction with Titanic train dataset.


```{r}
#install.packages("titanic")
library(titanic)
set.seed(12233)
df = tibble(titanic_train)  #load dataset
```

### a. 

The package `titanic` has data on the survival information of Titanic passengers including `Name`. Extract titles from the names of the passengers. These include `mr`, `mrs`, etc. Use suitable regex that consists of a look ahead pattern that tells us to look for something followed by a space, one or more instances of character/s, and a look behind to match patterns to a period after the string, respectively.


```{r}
# Define the regex
# reg <- 
```


### b. 

Plot the distribution of the the titles using either `geom_col()` or `geom_bar()` after properly reordering the variable based on their counts.



### c. 

Extract family names from `Name` and plot the 10 most frequent family names. Remember to extract the pattern to include any kinds of characters (.) that are 1 character or more in length (+) and are followed by a comma(?=``\\``,). What is/are the most popular last name/names?

```{r}
#reg <- 
```



### d. 

Write a function that plots the top 10 last names of passengers beginning and ending with certain letters passed as an argument to the function. Pass the `data tibble`, `letter corresponding to the beginning of the last name`, and `letter corresponding to the ending of the last name` as arguments to the function, and a plot (a distribution plot depicting the last names and their count) as the output of the function.


