---
title: "Homework 2"
output: pdf_document
---

## Name: Put your name here

## I worked with: 

**Click the "Knit" button in RStudio to knit this file to a pdf.**

--------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, 
                      warning = FALSE, message = FALSE)
# add packages here 
```



## Problem 1: Spot the error

*answer:* write your answer here

```{r}
# put your code here
```


----------------------------------

## Problem 2: Penguins

### a. 

*answer:* write your answer here

```{r}
# put your code here
```

### b. 

*answer:* write your answer here

```{r}
# put your code here
```

### c. 

*answer:* write your answer here

```{r}
# put your code here
```


### d. 

*answer:* write your answer here

```{r}
# put your code here
```

----------------------------------

## Problem 3: Storm paths by year

### a. 

*answer:* write your answer here

```{r}
# put your code here
```


### b. 

*answer:* write your answer here

```{r}
# put your code here
```



-------------------------

## Problem 4:  explain command (no R needed)


### a. 

*answer:* write your answer here and your Markdown table below

### b. 

*answer:* write your answer here and your Markdown table below

### c. 

*answer:* write your answer here and your Markdown table below

### d. 

*answer:* write your answer here and your Markdown table below

### e. 

*answer:* write your answer here and your Markdown table below


