---
title: "Individual homework 2"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Roles for this assignment:

## Please push your assignment to GitHub by 10:00pm (Central) Monday, Apr 8. 

You are currently in the GitHub repository (repo) for `hw2-username`. The assignment prompt is shown below (i.e. in `README.Rmd`). You can view this online in your homework 2 GitHub repository as a Markdown file(`README.md`) or a pdf.

Please **use `hw2.Rmd` to complete this assignment**. Be sure to **knit your file to PDF before your final push to GitHub**.

## Homework process

For help on the homework process, review

-   [Assignments in Stat 220](https://stat220-spring24.netlify.app/assignments) for
    content/formatting questions.

-   [GitHub Guide for Students in Stat 220](https://stat220-spring24.netlify.app/github_tutorial)
    for Git and Github instructions.

When you are done with your homework, **don’t forget to push your changes to ALL files back to GitHub!** This means you should commit and push all related files, not just your final PDF. Additionally, ensure you post the link to your **GitHub repository to Gradescope for the final submission and grading**. This step is crucial as it allows for a comprehensive review of both your code and the rendered output.


--------------------------------------------------------------

## Assignment prompt


## Problem 1: Spot the error

Explain why the following command does not color the data points blue, then write down the command that will turn the points blue.

```{r}
library(ggplot2)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```



## Problem 2: Penguins 

Load the data `penguins` using the command

```{r}
# install.packages("palmerpenguins")  # run once then DELETE FROM RMD
data(penguins, package = 'palmerpenguins')
```

Look at help file for for variable info: `?palmerpenguins::penguins`. 

### a. 

Create a scatterplot of `body_mass_g` (y) against `bill_length_mm` (x) and separate the plot into facets by `species`. 

### b. 

Add regression lines to the plot in (a) for each species with no standard error shading. See `?geom_smooth` to see how to add a regression (`lm`) line instead of a wigglely smoother. 

### c. 

Instead of faceting by `species`, use `species` to color the points  and regression lines. Use a color scheme for the points that is not the default coloring. 

### d. 

Which visual, (b) or (c), makes it easiest to compare the **slopes** of the lines for each species? Explain.

----------------------------------------------

## Problem 3: Storm paths by year

Install the package `nasaweather`, if needed, and load the `storms` data.

```{r}
# install.packages("nasaweather")
data(storms, package = "nasaweather")
```

### a. 

Use `geom_path()` to plot the path of each tropical storm in this data set where you 

- use color to distinguish the storms from one another
- use faceting to plot each `year` on its own panel
- change your coordinate system to a map using `coord_map()`.

Hint: To use color to distinguish storms but not include the legend of storm names/colors you can add `scale_color_discrete(guide="none")`

### b. 

Add a map of the US, Mexico, Canada and the UK to your map in (a) so you can get a better idea where these storms are traveling, truncating the longitude as suggested in the hints.

Hints:

- Look at the help file for `?map_data`. To create map data with only the counties around the storm paths, use data from the the `world` map from `map_data` with the following `region`s specified: `usa`, `mexico`, `canada` and `uk`.

- Add the map data for the regions of interest using `geom_polygon`. (Hint: make sure your paths map doesn't have `color` as a global `aes`.)

- Modify your `coord_map` from part (a) to add the arguments `xlim`  and `ylim` to set the min and max limits of your graph to the min/max latitude and longitude of the `storms`.

------------------------------------------


## Problem 4: explain command (no R needed)

Consider the data set shown in the table below. Each row represents a class and `classType` of `S` denotes a statistics class and `C` denotes a CS class. The variable `m` counts the number of mac users in the class and `w` counts the number of windows users. 

classType | m | w 
------ | -- | --
C | 10 | 4
C | 3 | 1
C | 7 | 3
S | 2 | 7
S | 7 | 10

What data set will be produced by the following commands? Describe (**in words**) how the original data set is being modified and show what it looks like using an R Markdown table (like the one used above) to display the new data set. Assume the original data set is named `mydata`. (No credit will be given creating the fake data frame and just  running the code chunks.)

### a.
```{r, eval=FALSE}
mydata %>% 
  filter(classType == "C") %>%
  select(m, w)
```

### b. 
```{r, eval=FALSE}
mydata %>% 
  mutate(ratioW = w/sum(w))
```

### c. 
```{r, eval=FALSE}
mydata %>% 
  group_by(classType) %>%
  mutate(ratioW = w/sum(w))
```

### d. 
```{r, eval=FALSE}
mydata %>% 
  group_by(classType) %>%
  summarize(Y = sum(w+m))
```

### e. 
```{r, eval=FALSE}
mydata %>% 
  group_by(classType) %>%
  mutate(X = w+m, Y = sum(w+m))
```

---------------------------------------------------------------------

