---
title: "Class Activity 6"
author: "Your name here"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
format:
  pdf: default
  html:
    df_print: paged
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
  error: true
---

```{r setup, include=FALSE}
# load the necessary libraries
library(dplyr)
library(ggplot2)
library(babynames)
```


We will work with the `babynames` dataset again in this class activity. The header of the dataset looks like this:

```{r}
knitr::kable(head(babynames))
```

In this tutorial, we will learn about the five main verbs of dplyr and how to use them to manipulate data:

- `select()`: Choose columns from a data frame
- `filter()`: Choose rows based on a condition
- `arrange()`: Sort the rows of a data frame
- `mutate()`: Add new columns based on existing columns
- `summarise()`: Aggregate data and compute summary statistics

## Problem 1: `select()`

Which of these is NOT a way to select the `name` and `n` columns together?

```{r, results='hide'}
select(babynames, -c(year, sex, prop))  #1
select(babynames, name:n)   #2
select(babynames, starts_with("n")) #3
select(babynames, ends_with("n"))  #4
```

*Answer:* 

\vspace*{1in}


## Problem 2: `filter()`

Use `filter()` with the logical operators to extract:

### a. All of the names where prop is greater than or equal to 0.08

```{r}

```

### b. All of the babies named "Margaret"

```{r}

```


### c. Use `filter()` to choose all rows where name is "John" and sex is "M".


```{r}

```


\vspace*{1in}

## Problem 3: `arrange()`

### a. Use `arrange()` to sort the babynames dataset by the prop column in descending order.

```{r}

```


### b. Use `arrange()` to sort the babynames dataset by year (ascending) and then by prop (descending).

```{r}

```


\vspace*{1in}


## Problem 4: `mutate()`

### a. Use `mutate()` to create a new column called decade which contains the decade the record is in (e.g., 1990 for the years 1990-1999).

```{r}

```


\vspace*{1in}


## Problem 5: `summarize()` or `summarise()`

Use the codes mentioned so far to compute three statistics:

- the total number of children who ever had your name
- the maximum number of children given your name in a single year
- the mean number of children given your name per year


```{r}



```


\vspace*{1in}


## Problem 6

### a. Use `min_rank()` and `mutate()` to rank each row in babynames from largest prop to smallest prop.

```{r}

```


### b. Compute each name's rank within its year and sex.

```{r}

```


### c. *(Optional)* Then compute the median rank for each combination of name and sex, and arrange the results from highest median rank to lowest.


```{r}

```


