---
title: "Class Activity 22"
author: "Your name here"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      size = "small", 
                      collapse = TRUE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      error = TRUE) # change it to TRUE

# load the necessary libraries
library(tidyverse) 
library(tidymodels)
library(mlbench)     # for PimaIndiansDiabetes2 dataset
library(janitor)
library(parsnip)
library(kknn)
library(ggthemes)
library(purrr)
library(forcats)

select <- dplyr::select
```



## Group Activity 1

Load the `mlbench` package to get `PimaIndiansDiabetes2` dataset.

```{r}
# Load the data - diabetes
data(PimaIndiansDiabetes2)
db <- PimaIndiansDiabetes2
db <- db %>% drop_na() %>% mutate(diabetes = fct_rev(factor(diabetes))) 
db_raw <- db %>% select(glucose, insulin, diabetes)
```


a. Split the data `75-25` into training and test set using the following code. 

```{r}
set.seed(123)


```
  
  
b. Follow the steps to train a 7-NN classifier using the `tidymodels` toolkit 
    

```{r}
# define recipe and preprocess the data
db_recipe <- 
```
    

```{r}
# specify the model
db_knn_spec7 <- 
```


```{r}
# define the workflow
db_workflow <-
```


```{r}
# fit the model on the training data
db_fit <- 
```
    
    
c. Classify the penguins in the `test` data frame. 

```{r}
test_features <- db_test %>% select(glucose, insulin) 

# predict
db_pred <- 

  
# combine the results

```


-----------------------------------------------------------------------------------------------

## Group Activity 2

Calculate the accuracy, sensitivity, specificity, and positive predictive value (PPV) using the following confusion matrix.

```{r}
conf_mat(db_results, truth = diabetes, estimate = predicted)
```


```{r}
# your r-code

```


## Extra: Code to recreate the plot in the `slides` for the `diabetes` dataset.


```{r}
metrics_for_k <- function(k, db_train, db_test){
db_knn_spec <- nearest_neighbor(mode = "classification",
                             engine = "kknn",
                             weight_func = "rectangular",
                             neighbors = k)

db_knn_wkflow <- workflow() %>%
  add_recipe(db_recipe) %>%
  add_model(db_knn_spec)

db_knn_fit <- fit(db_knn_wkflow, data = db_train)
test_features <- db_test %>% select(glucose, insulin)
nn1_pred <- predict(db_knn_fit, test_features, type = "raw")

db_results <- db_test %>% 
  select(diabetes) %>% 
  bind_cols(predicted = nn1_pred)
custom_metrics <- metric_set(accuracy, sens, spec, ppv)

metrics <- custom_metrics(db_results,
               truth = diabetes,
               estimate = predicted) 
metrics <- metrics %>% select(-.estimator) %>% mutate(k = rep(k,4))

return(list = metrics)
}



# run the function and store it as "optim.results"




# plot

optim.results %>%
  ggplot(aes(x = , y = , color = ))) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  theme_minimal() +
  ggthemes::scale_color_wsj() + 
  scale_x_continuous(breaks = k) +
  theme(panel.grid.minor.x = element_blank(),
        axis.text=element_text(size=6, angle = 20))+
  labs(color='Metric', y = "Estimate", x = "K") 
```


