---
title: "Class Activity 10"
author: "Your name here"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      size = "small", 
                      collapse = TRUE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      error = TRUE)

# load the necessary libraries
library(tidyverse)
library(janitor)
library(tidyr)
library(lubridate)
library(readr)
```


\vspace*{1in}

## Your Turn 1


```{r}
students <- tibble(
  id = 1:24,
  grade = sample(c("9th", "10th", "11th"), 24, replace = TRUE),
  region = sample(c("North America", "Europe", "Asia", 
                    "South America", "Middle East", "Africa"), 24, replace = TRUE),
  score = round(runif(24,50, 100))
)
```


### a. Create a new column `grade_fac` by converting the grade column into a factor. Reorder the levels of `grade_fac` to be "9th", "10th", and "11th". Sort the dataset based on the `grade_fac` column.

*Answer:* 

```{r}
students_a <- students %>%
  mutate(grade_fac = factor(grade)) %>%
  mutate(grade_fac = fct_relevel(grade_fac, c("9th", "10th", "11th"))) %>%
  arrange(grade_fac) 
print(students_a, n = 24)
```


### b. Create a new column `region_fac` by converting the `region` column into a factor. Collapse the `region_fac` levels into two categories: "Male" and "Female". Count the number of students in each collapsed region category.

```{r}
students_b <- students_a %>%
  mutate(region_fac = factor(region)) %>%
  mutate(region_collapsed = fct_collapse(region_fac, 
                                         Americas = c("North America", "South America"), 
                                         EMEA = c("Europe", "Middle East", "Africa"), 
                                         Asia = "Asia")) %>%
  count(region_collapsed)
print(students_b)
```


### c. Create a new column `grade_infreq` that is a copy of the `grade_fac` column. Reorder the levels of `grade_infreq` based on their frequency in the dataset. Print the levels of `grade_infreq` to check the ordering.

```{r}
students_c <- students_a %>%
  mutate(grade_infreq = grade_fac) %>%
  mutate(grade_infreq = fct_infreq(grade_infreq))

levels(students_c$grade_infreq)
```


### d. Create a new column `grade_lumped` by lumping the least frequent level of the `grade_fac` column into an 'Others' category.
Count the number of students in each of the categories of the `grade_lumped` column.



```{r}
students_d <- students_a %>%
  mutate(grade_lumped = fct_lump(grade_fac, n = 1, other_level = "Others")) %>%
  count(grade_lumped)
students_d
```



## Your Turn 2

Lets import the `gss_cat` dataset from the `forcats` library. This datast contains a sample of categorical variables from the General Social survey.

```{r}
# import gss_cat dataset from forcats library
forcats::gss_cat
```

Use `gss_cat` to answer the following questions.

### a. Which religions watch the least TV?

```{r}
# your r-code

gss_cat %>%
  drop_na(tvhours) %>%
  group_by(relig) %>%
  summarize(tvhours = mean(tvhours)) %>%
  ggplot(aes(tvhours, fct_reorder(relig, tvhours))) +
    geom_point()
```



### b. Do married people watch more or less TV than single people?

```{r}
# your r-code

gss_cat %>%
  drop_na(tvhours) %>%
  group_by(marital) %>%
  summarize(tvhours = mean(tvhours)) %>%
  ggplot(aes(tvhours, fct_reorder(marital, tvhours))) +
    geom_point()

```


c. Collapse the `marital` variable to have levels `Married`, `not_married`, and `No answer` .Include `"Never married"`, `"Divorced"`, and "`Widowed"` in `not_married`


```{r}
# your r-code

gss_cat %>%
  drop_na(tvhours) %>%
  select(marital, tvhours) %>%
  mutate(
    maritalStatus =
      fct_collapse(
        marital,
        Married = c("Married",
                         "Separated"),
        not_married = c("Never married",
                    "Divorced",
                    "Widowed"))
  )

```


