---
title: "Class Activity 10"
author: "Your name here"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
format:
  pdf: default
  html:
    df_print: paged
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
  error: true
---

```{r setup, include=FALSE}
# load the necessary libraries
library(tidyverse)
library(tidyr)
library(forcats)
```

\vspace*{1in}

## Your Turn 1

```{r}
students <- tibble(
  id = 1:24,
  grade = sample(c("9th", "10th", "11th"), 24, replace = TRUE),
  region = sample(c("North America", "Europe", "Asia", "South America", "Middle East", "Africa"), 24, replace = TRUE),
  score = round(runif(24,50, 100))
)
```

### a. Create a new column `grade_fac` by converting the grade column into a factor. Reorder the levels of `grade_fac` to be "9th", "10th", and "11th". Sort the dataset based on the `grade_fac` column.

*Answer:*

```{r}
students_a <- 
```

### b. Create a new column `region_fac` by converting the `region` column into a factor. Collapse the `region_fac` levels into three categories: "Americas", "EMEA" and "Asia". Count the number of students in each collapsed region category.

```{r}
students_b <- students_a 
```

### c. Create a new column `grade_infreq` that is a copy of the `grade_fac` column. Reorder the levels of `grade_infreq` based on their frequency in the dataset. Print the levels of `grade_infreq` to check the ordering.

```{r}
students_c <- students_a 
```

### d. Create a new column `grade_lumped` by lumping the least frequent level of the `grade_fac` column into an 'Others' category.

Count the number of students in each of the categories of the `grade_lumped` column.

```{r}
students_d <- 
```

## Your Turn 2

Lets import the `gss_cat` dataset from the `forcats` library. This dataset contains a sample of categorical variables from the General Social survey.

```{r}
# import gss_cat dataset from forcats library
forcats::gss_cat
```

Use `gss_cat` to answer the following questions.

### a. Which religions watch the least TV?

```{r}
# your r-code

gss_cat %>%
 
```

### b. Do married people watch more or less TV than single people?

```{r}
# your r-code

gss_cat %>%

```

c.  Collapse the `marital` variable to have levels `Married`, `Not_married`, and `No_answer` . Include `"Never married"`, `"Divorced"`, and "`Widowed"` in `Not_married`

```{r}
# your r-code

gss_cat %>%
  

```
