---
title: "Class Activity 24"
author: "Your name here"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      size = "small", 
                      collapse = TRUE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      error = TRUE) # change it to TRUE

# load the necessary libraries
library(tidyverse) 
library(ggthemes)
library(janitor)
library(broom)
library(mlbench)
library(tidymodels)
library(probably)

select <- dplyr::select
theme_set(theme_stata(base_size = 10))

data(PimaIndiansDiabetes2)
db <- PimaIndiansDiabetes2
db <- db %>% drop_na()  %>% 
  mutate(diabetes = fct_relevel(diabetes, c("neg", "pos"))) # Relevels 'diabetes' factor to ensure 'neg' comes before 'pos'
```


\vspace*{1in}


## Group Activity 1

In this activity, we will calculate the probability of diabetes for a glucose level of 150 mg/dL using the logistic regression coefficients $\beta_0 = -5.61$ and $\beta_1 = 0.0392$.

### a. Calculate Log Odds

First, calculate the log odds for a glucose level of 150 mg/dL.

```{r}

```


### b. Convert Log Odds to Odds

```{r}

```


### c. Convert Odds to Probability

Finally, convert the odds to probability.

```{r}

```

The probability of having diabetes at a glucose level of 150 mg/dL is calculated to be 

---------------------------------------------------------

## Group Activity 2

a. Let's fit the logistic regression model to predict the `diabetes` status given the `glucose` level.

```{r}
set.seed(12345)
db_single <- db %>% select(_____, ______)
db_split <- initial_split(db_single, prop = 0.80)

# Create training data
db_train <- db_split %>% training()

# Create testing data
db_test <- db_split %>% testing()

fitted_logistic_model <- logistic_reg() %>% # Call the model function
        # Set the engine/family of the model
        set_engine() %>%
        # Set the mode
        set_mode() %>%
        # Fit the model
        fit()

tidy(fitted_logistic_model)
```


b. We are interested in predicting the `diabetes` status of patients depending on the amount of `glucose`. Verify that the `glucose` value of 143.11 gives the probability of having `diabetes` as 1/2.


$$log\left(\frac{p}{1-p}\right)  = \beta_0 + \beta_1x$$
*Answer:*



c. What value of glucose is needed to have a probability of diabetes of 0.75?

*Answer:*



d. Make a classifier that classifies the diabetes status of new patients with a threshold of 0.5, i.e, a new patient is classified as negative if the estimated class probability is less than 0.75. Also, create a confusion matrix of the resulting predictions. Evaluate the model based on accuracy, sensitivity, specificity, and ppv.


```{r}
# Prediction Probabilities

pred_prob <- predict(,  new_data = ,   type = )

db_results <- db_test %>% bind_cols(pred_prob) %>%
  mutate(.pred_class = make_two_class_pred(.pred_neg, levels(diabetes), threshold = .55)) %>%
  select(diabetes, glucose, contains(".pred"))


db_results %>%  
  conf_mat(diabetes,.pred_class) %>% 
  autoplot(type = "heatmap")

# Evaluating the model
eval_metrics <- metric_set(accuracy, sensitivity, specificity, ppv)

eval_metrics(data = db_results,
             truth = diabetes,
             estimate = .pred_class) %>% select(-2) 
```


e. Generate a ROC Curve and Determine the Optimal Threshold: Evaluate the performance of your diabetes prediction model by plotting a ROC curve. Use the curve to identify the point that is closest to the top-left corner (maximizing sensitivity and minimizing 1 - specificity), and back-calculate to find the corresponding optimal threshold. This threshold represents the best balance between sensitivity (true positive rate) and specificity (false positive rate). 


```{r}
library(yardstick)

# Predict probabilities on the test set
diabetes_prob <- predict(fitted_logistic_model, db_test, type = "prob")
diabetes_results <- db_test %>% 
  select(diabetes) %>%
  bind_cols(diabetes_prob)

# Generate ROC curve data
roc_data <- roc_curve()


# Plot the ROC curve
roc_plot <- ggplot(roc_data, aes(x = , y = )) +
  geom_line() +
  geom_abline(linetype = "dashed", color = "gray", slope = 1) +
  theme_minimal() +
  labs(title = "ROC Curve", x = "1 - Specificity (False Positive Rate)", y = "Sensitivity (True Positive Rate)")
roc_plot

# Identify the closest point to the top-left corner
optimal_point <- roc_data %>%
  mutate(distance = sqrt((1 - specificity)^2 + (sensitivity - 1)^2)) %>%
  arrange(distance) %>%
  slice(1)
optimal_point
```


