---
title: "Individual Homework 1"
output: pdf_document
---

## Name: Put your name here

## I worked with: 

**Click the "Knit" button in RStudio to knit this file to a pdf.**

--------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Problem 1: Markdown output


### a. 

*answer:* write your answer here



### b.

*answer:* write your answer here



---------------------------

## Problem 2: Logical vectors 

### a.

*answer:* write your answer here

```{r}
# put your code here
```


### b. 

*answer:* write your answer here

```{r}
# put your code here
```

### c. 

*answer:* write your answer here

```{r}
# put your code here
```

-------------------------------------------

## Problem 3: Coercion

### a.

*answer:* write your answer here



### b. 

*answer:* write your answer here



### c. 

*answer:* write your answer here



### d.

*answer:* write your answer here



### e.

*answer:* write your answer here


-----------------------------------------------

## Problem 4: Data frames


### a. 

*answer:* write your answer here

```{r}
# put your code here
```

### b. 

*answer:* write your answer here

```{r}
# put your code here
```

### c. 

*answer:* write your answer here

```{r}
# put your code here
```

### d. 

*answer:* write your answer here

```{r}
# put your code here
```

--------------------------

## Problem 5: Lists


### a. 

*answer:* write your answer here

```{r}
# put your code here
```

### b. 

*answer:* write your answer here

```{r}
# put your code here
```

### c. 

*answer:* write your answer here

```{r}
# put your code here
```

### d. 

*answer:* write your answer here

```{r}
# put your code here
```

---------------------------------

## Problem 6: More lists

### a. 


*answer:* write your answer here

```{r}
# put your code here
```


### b. 

*answer:* write your answer here

```{r}
# put your code here
```

### c. 
`
*answer:* write your answer here

```{r}
# put your code here
```

