---
title: "Class Activity 11"
author: "Your name here"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      size = "small", 
                      collapse = TRUE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      error = TRUE)

# load the necessary libraries
library(tidyverse)
library(stringr)
```


\vspace*{1in}


## Problem 1

Let's learn about combining strings with different separators first.

```{r}
place <- "Central Park"
activity <- "jogging"
activities <- c("jogging", "picnicking", "boating")
my_sentence <- str_c(place, " is great for ", activity, ".", sep = "")
my_sentence
```


a. What happens when a `str_c` entry is a vector?

*Answer:* When an entry in `str_c` is a vector, it will combine the strings with each element of the vector, creating multiple combined strings.

```{r}
my_sentences <- str_c(place, " is great for ", activities, ".", sep = "")
my_sentences
```


b. How do you combine strings with `str_glue`?

*Answer:* You can combine strings with `str_glue` using curly braces `{}` to insert variables directly into the string.


```{r}
my_sentence <- str_glue("{place} is great for {activity}.")
my_sentence

my_sentences1 <- str_glue("{place} is great for {activities}.")
my_sentences1
```


c. What does `str_flatten` do?

*Answer:* `str_flatten` collapses a character vector into a single string by concatenating the elements with a specified separator.


```{r}
p <- str_flatten(my_sentences, collapse = " and ")
writeLines(p)
```


d. What will using a `\n` separator do in the command below?

*Answer:* Using a `\n` separator in the command will insert a newline character between the strings being combined, making them display on separate lines when printed.


```{r}
p <- str_c(place, " is great for ", activity, sep = "\n")
writeLines(p)
```


e. Does `str_length` count spaces and special characters??

*Answer:* Yes, `str_length` counts spaces and special characters as part of the string's length.


```{r}
p
str_length(p)
```


f. How do you count the number of `e`'s in a string?

*Answer:* You can count the number of `e`'s in a string using `str_count` with a pattern that matches the character 'e'.


```{r}
text <- "The quick brown fox jumps over the lazy dog."
pattern <- "e"
vowel_count <- str_count(text, pattern)
vowel_count
```



g. What happens with negative positions?

*Answer:* Negative positions in `str_sub` count the positions from the end of the string rather than from the beginning.


```{r}
str_sub(my_sentence, start = -3, end = -1)
```


h. How do you extract a `substring` with positive and negative positions?

*Answer:* You can extract a `substring` with positive and negative positions using `str_sub` and specifying the start and end positions with either positive or negative numbers.


```{r}
my_sentence <- "Central Park is great for jogging."
positive_substr <- str_sub(my_sentence, start = 1, end = 12)
negative_substr <- str_sub(my_sentence, start = -8, end = -1)
positive_substr
negative_substr
```


i. With a vector of positions?

*Answer:* Using a vector of positions with `str_sub` will extract `substrings` starting and ending at the specified positions in the vector.


```{r}
str_sub(my_sentence, start = c(1, 9), end = c(4, 15))
```


j. How do you extract multiple `substrings` using a vector of positions?

*Answer:* You can extract multiple `substrings` using a vector of positions with `str_sub` by specifying the start and end positions in separate vectors.


```{r}
my_sentence <- "Central Park is great for jogging."
substrs <- str_sub(my_sentence, start = c(1, 14, 24), end = c(12, 19, 30))
substrs
```


-----------------------------------------------------------------------

## Problem 2

a. Use the string parsing functions that you learned today to do tasks described in the comments below: 

```{r}
s1 <- "12%"  # remove %
s2 <- "New Jersey_*"  # remove _*
s3 <- "2,150"    # remove comma(,)
s4 <- "Learning #datascience is fun!"   # extract #datascience
s5 <- "123 Main St, Springfield, MA, 01101"   # separate info

# Cleaning steps
s1_clean <- str_replace(s1, "%", "")
s2_clean <- str_replace(s2, "_\\*", "")
s3_clean <- str_replace(s3, ",", "")
s4_clean <- str_extract(s4, "#\\w+")
s5_clean <- str_split(s5, ",\\s?")

# Print cleaned strings
s1_clean
s2_clean
s3_clean
s4_clean
s5_clean
```


## Problem 3

a. Use the string parsing functions that you learned today to do tasks described in the comments below: 

```{r}
s1 <- "25%"  # remove %
s2 <- "Los Angeles_#"  # remove _#
s3 <- "1,250"    # remove comma(,)
s4 <- "Discover #machinelearning today!"   # extract #machinelearning
s5 <- "456 Main St, San Francisco, CA, 94107"   # separate info

# Cleaning steps
s1_clean <- str_replace(s1, "%", "")
s2_clean <- str_replace(s2, "_#", "")
s3_clean <- str_replace(s3, ",", "")
s4_clean <- str_extract(s4, "#\\w+")
s5_clean <- str_split(s5, ",\\s?")

# Print cleaned strings
s1_clean
s2_clean
s3_clean
s4_clean
s5_clean
```


## Problem 4

a. Let's look at the following dataset containing information about movies and their release years. We'll extract the release year from the movie title, create a new column with decades, and count the number of movies in each decade.


```{r}
# Sample dataset
movies <- tibble(
  title = c(
    "The Godfather (1972)", "Pulp Fiction (1994)", "The Dark Knight (2008)",
    "Forrest Gump (1994)", "The Shawshank Redemption (1994)", "The Matrix (1999)",
    "Inception (2010)", "Interstellar (2014)", "Parasite (2019)", "Fight Club (1999)"
  )
)
movies
```


```{r}
# Processing the dataset
movies_processed <- movies %>%
  mutate(
    release_year = as.integer(str_extract(title, "\\d{4}")),
    decade = floor(release_year / 10) * 10
  ) %>%
  count(decade) %>%
  rename(num_movies = n)

# Print the processed dataset
movies_processed
```


