---
title: "Midterm II Study Guide and Review"
author: "Deepak Bastola"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---


## Midterm II Study Guide

Format: In-class with open-ended questions.

One-sided Cheat-sheet allowed (A4 paper) and a basic calculator allowed.

  - `shiny` and `purrr` cheatsheet will be provided
  - You are not permitted to use a laptop or classroom computer. 

### Topics

- The exam covers iterations, functionals, web scraping, shiny, kNN, linear regression and logistic regression (through Mon. 05/20)

- You will be evaluated on your understanding of the tools, algorithms, Shiny function logic, and the construction of the workflow in R that we have covered in class. While I won't require you to write complex code from scratch, you should be ready to write brief segments of code. Other methods I might use to assess your knowledge of R include:

  - Identifying the error in written code.
  - Putting lines of code in order to complete a specified task.
  - Describing the output resulting from a code/code-chunk.


```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, warning = FALSE, 
                      message=FALSE, include=TRUE, 
                      fig.height = 4, fig.width = 5)


library(tidyverse) 
library(tidymodels)
library(mlbench)     # for PimaIndiansDiabetes2 dataset
library(yardstick) # extra package for getting metrics
library(parsnip) # tidy interface to models
library(ggthemes)
library(vip)
library(ISLR)
library(rpart.plot)
library(janitor)
library(ranger)
library(lubridate)
# prepare datasets

library(datasets)
mydata <- as.data.frame(matrix(ldeaths, nrow = 6))
colnames(mydata) <- str_to_lower(month(1:12, label = TRUE))
mydata <- mydata %>% tibble() %>% mutate(year = seq(1974, 1979)) %>% relocate(year, .before = jan)
```



\newpage

# Sample Questions


## Q1 

Given below are the monthly deaths from bronchitis, emphysema and asthma in the UK from 1974 to 1979.

```{r}
knitr::kable(mydata)
```



### a.  Describe what is returned by the code below, including the type of R object produced, the length or dimension of the object, and the information contained in the object.

```{r, eval=FALSE}
map_dbl(mydata %>% select(-1), mean) %>% mean()
```

\vspace*{1in}


### b. Describe what is returned by the code below, including the type of R object produced, the length or dimension of the object, and the information contained in the object.

```{r, eval=FALSE}
ratio_fun <- function(x) quantile(x, probs= c(0.25, 0.5, 0.75))
map_dfc(mydata %>% select(-1), ratio_fun)
```


\vspace*{1.5in}


### c. Describe what is returned by the code below, including the type of R object produced, the length or dimension of the object, and the information contained in the object.

```{r, eval=FALSE}
lapply(mydata %>% select(-1), ratio_fun) %>%  
  unlist()
```


\newpage

## Q2 Shiny UI and Server Logic

### a. Describe the relationship between `ui` and `server` in a Shiny application. How do they interact?

\vspace*{1.5in}

### b. Given the code snippet below, identify and explain any errors that would prevent the app from displaying a histogram of the generated data.

```{r, eval=FALSE}
library(shiny)
ui <- fluidPage(
  titlePanel("Data Histogram"),
  sidebarLayout(
    sidebarPanel(
      sliderInput("obs", "Number of observations:", min = 0, max = 1000, value = 500)
    ),
    mainPanel(
      plotOutput("distPlot")
    )
  )
)

server <- function(input, output) {
  output$distPlot <- renderPlot({
    data <- rnorm(input$obs, input = 100)  
    hist(data)
  })
}

shinyApp(ui = ui, server = server)
```

\vspace*{1.5in}

### c. Consider a Shiny app that allows users to select a dataset from a dropdown menu and displays a summary of that dataset. Identify potential problems in this hypothetical Shiny code snippet and suggest improvements.

```{r, eval=FALSE}
library(shiny)
ui <- fluidPage(
  selectInput("dataset", "Choose a dataset:", choices = c("mtcars", "iris")),
  tableOutput("summaryTable")
)

server <- function(input, output) {
  data <- reactive({
    switch(input$dataset,
           "mtcars" = mtcars,
           "iris" = iris)
  })

  output$summaryTable <- renderTable({
    summary(data())
  })
}

shinyApp(ui = ui, server = server)

```

\vspace*{1.5in}

### d. Given a Shiny application that uses actionButton to trigger calculations, explain the purpose of this input type and describe a scenario where it might be useful.

```{r, eval=FALSE}
ui <- fluidPage(
  actionButton("calc", "Calculate"),
  numericInput("num", "Enter a number:", 10),
  textOutput("result")
)

server <- function(input, output) {
  observeEvent(input$calc, {
    result <- sqrt(input$num)
    output$result <- renderText({
      paste("The square root is:", result)
    })
  })
}

shinyApp(ui = ui, server = server)

```


\newpage


## Q3 Web scraping

```html
<!DOCTYPE html>
<html lang="en">
<head>
    <meta charset="UTF-8">
    <title>Sample Web Page</title>
</head>
<body>
    <h1>Welcome to My Website</h1>
    <p>This is a paragraph of text on the main page.</p>
    <a href="/about">About Us</a>
    <a href="/contact">Contact Us</a>
</body>
</html>
```

Assume the HTML above is served from `http://mywebpage.com.`


### a. Write the R code to extract and print the text inside the `<h1>` tag using `rvest` and/or `polite.`


```{r, eval=FALSE, echo=FALSE}
library(rvest)
html_code <- read_html('http://mywebpage.com')
h1_text <- html_code %>% html_node('h1') %>% html_text()
print(h1_text)
```

\vspace*{1in}

### b. Write the R code to extract the href attributes of all links in the document.

```{r, eval=FALSE, echo=FALSE}
links <- html_code %>% html_nodes('a') %>% html_attr('href')
print(links)
```

\vspace*{1in}

### c. Given the base URL `http://mywebpage.com`, write the R code to convert the relative URLs found in the href attributes into absolute URLs.

```{r, eval=FALSE, echo=FALSE}
absolute_links <- url_absolute(links, 'http://example.com')
print(absolute_links)
```

\vspace*{1in}

### d. Write the R code to extract and print the text of each link (<a> tag).

```{r, eval=FALSE, echo=FALSE}
link_texts <- html_code %>% html_nodes('a') %>% html_text()
print(link_texts)
```

\newpage

## Q4 : K-nearest neighbor

Consider the `Smarket` dataset, which provides daily percentage returns for the S&P 500 stock index from 2001 to 2005. Your task is to fit a K-nearest neighbor model to predict the direction of stock returns.

#### a. . The following code splits the Smarket dataset into a training set and a test set. Explain the purpose of splitting the data and what stratification accomplishes in this context.

```{r}
set.seed(1234)

data_Smarket <- as_tibble(Smarket)
split <- initial_split(data_Smarket, strata = Direction, prop = 4/5)
Smarket_train <- training(split)
Smarket_test <- testing(split)

# glimpses of data
glimpse(Smarket_train)
glimpse(Smarket_test)
```





#### b. The following trained model is used to produce a data-frame of the actual and predicted `Direction` in the test dataset. Call this data-frame `Smarket_results`. What information does `Smarket_results` contain? What is the dimension of this dataset? Explain.



```{r}
Smarket_recipe <- recipe(Direction ~ Lag1 + Lag2 + Lag3 + Year + Volume, 
                         data = Smarket_train) %>% 
  step_center(all_predictors()) %>% 
  step_scale(all_predictors()) 

Smarket_knn_spec <- nearest_neighbor(mode = "classification",
                             engine = "kknn",
                             weight_func = "rectangular",
                             neighbors = 5)

Smarket_workflow <- workflow() %>% 
  add_recipe(Smarket_recipe) %>%
  add_model(Smarket_knn_spec)

Smarket_fit <- fit(Smarket_workflow, data = Smarket_train)

test_features <- Smarket_test %>% select(Direction, Lag1, Lag2, Lag3, Year, Volume)
nn1_pred3 <- predict(Smarket_fit, test_features, type = "raw")
Smarket_results <- Smarket_test %>% 
  select(Direction) %>% 
  bind_cols(predicted = nn1_pred3) %>% mutate(Direction = as.factor(Direction))
```


```{r, eval = FALSE, echo=FALSE}
glimpse(Smarket_results)
```

\vspace*{2.5in}



#### c.  Calculate the sensitivity, specificity, accuracy, and positive predictive value of the classifier based on the confusion matrix below. Discuss the interpretation of these metrics in the context of this problem.


```{r}
conf_mat(Smarket_results, truth = Direction, estimate = predicted)
```

\newpage



## Q5 Miscellaneous

### (a) Differentiate between supervised and unsupervised learning by providing examples of situations where each would be most appropriate.


\vspace*{1.5in}


### (b) Discuss the significance of feature scaling in the K-NN algorithm and illustrate its impact with a hypothetical example.

\vspace*{1.5in}




### (c) (True/False) Recall is a more relevant metric than precision in scenarios where False Negatives are more detrimental than False Positives. Explain your answer.

\vspace*{1.5in}



### (d)  Explain the rationale behind data preprocessing in the KNN algorithm, discussing how the algorithm's performance might be influenced by unprocessed data.

\newpage



### (e) (Multiple Choice) Among the following values for K in K-NN, which is most likely to cause underfitting?

  1. 30
  2. 5
  3. 1
  
\vspace*{1.5in}

  
### (f) Logistic regression is a machine learning algorithm that is typically used to predict the probability of what kind of variable? Explain the reasoning behind your choice.

(A) categorical independent variable
(B) categorical dependent variable.
(C) numerical dependent variable.
(D) numerical independent variable.

\vspace*{1.5in}



### (g) For a K-NN classification model, how would you expect the model's performance to change as K increases? Discuss considering both bias-variance trade-off and the model's complexity.

\vspace*{1.5in}


### (h) In logistic regression, which link function is commonly used?

A) Logit
B) Probit
C) Identity
D) Log

\vspace*{1.5in}

### (i) Which method of data splitting ensures that the training and test sets have approximately the same percentage of samples of each class?

A) Random split
B) Stratified split
C) Sequential split
D) Clustered split




