---
title: "Midterm II"
date: "`r Sys.Date()`"
output:
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, warning = FALSE, 
                      message=FALSE, include=TRUE, 
                      fig.height = 4, fig.width = 5, error = TRUE)

library(tidyverse)
library(readr)
library(dplyr)
library(tidyr)
library(purrr)
library(forcats)
library(stringr)
library(lubridate)

# prepare datasets

library(datasets)
mydata <- as.data.frame(matrix(ldeaths, nrow = 6))
colnames(mydata) <- str_to_lower(month(1:12, label = TRUE))
mydata <- mydata %>% tibble() %>% mutate(year = seq(1974, 1979)) %>% relocate(year, .before = jan)
```


## Your name: 

\vspace*{1in}

# Questions



## Q1 

Given below are the monthly deaths from bronchitis, emphysema and asthma in the UK from 1974 to 1979.

```{r}
knitr::kable(mydata)
```


### a. 

Write a for loop that will return a vector of the ratio of the mean value to the median value for columns 2-13 in `mydata` (shown above).


```{r}
store_ratio <- rep(NA, 12)
for (i in 1:12){
  store_ratio[i] <- mean(mydata[[i+1]])/median(mydata[[i+1]])
}
store_ratio
```


### b.  

Describe what is returned by the code below, including the type of R object produced, the length or dimension of the object, and the information contained in the object.

```{r}
map_dbl(mydata %>% select(-1), mean) %>% mean()
```


### c. 

Describe what is returned by the code below, including the type of R object produced, the length or dimension of the object, and the information contained in the object.


```{r}
ratio_fun <- function(x) quantile(x, probs= c(0.25, 0.5, 0.75))
map_dfr(mydata %>% select(-1), ratio_fun, .id = "month")
```


### d. 

Describe what is returned by the code below, including the type of R object produced, the length or dimension of the object, and the information contained in the object.

```{r}
map_dfc(mydata %>% select(-1), ratio_fun)
```


### e. 

Describe what is returned by the code below, including the type of R object produced, the length or dimension of the object, and the information contained in the object.

```{r}
lapply(mydata %>% select(-1), ratio_fun) %>%  unlist()
```


### f.  

Describe what is done by the code below. What type of R object is `mydata_long`? What is the length or dimension of the `mydata_long`? 


```{r}
mydata_long <- mydata %>% pivot_longer(cols = 2:13,
                       names_to = "months",
                       values_to = "deaths") 
```



### g. 

Write a function that takes the data object from `f` and year as an argument and returns the median deaths for that particular year. Additionally, your function should give a warning if the user enters a year that  not in the dataset. 


```{r}
iqr_func <- function(data, year){
  if (year >= 1974 & year <= 1979){
    median(data %>% filter(year == year))
  } else {"Please enter a year from 1974 to 1979"}
}
iqr_func(mydata_long, 1975)
```



## Q2

### a. 

Consider the function below. Give the output produced by `myfun(3)` and explain how you arrived at your answer.

```{r}
myfun <- function(x) {
  if (x < 3 | x > 3){
  rep("hi", x)
  } else{
  rep("bye", x)
  }
}
```



## Q3 

Consider the following data frames for this question.

```{r}
df1 = data.frame(Id = c(1:6), Item = c(rep("Lake1", 3), rep("Lake2", 3)))
df2 = data.frame(Id = c(1, 3, 5), PH = c(rep("Acidic", 2), rep("basic", 1)))
df1
df2
```

What do the following codes do? Provide a sketch of the resulting data frame.

### a.


```{r}
semi_join(df1, df2)
```


### b.


```{r}
right_join(df1, df2)
```


### c.


```{r}
full_join(df1, df2)
```



## Q4

Consider the following string:

```{r}
my_string <- "A 5.1 magnitude earthquake strikes near San Jose, US Geological Survey 10 reports"
```


```{r}
str_subset(c("A", "5.1"), "[^\\d]")
```


### a.

Carefully explain what `str_view_all` will highlight in `x` for the given pattern.


```{r, eval=FALSE}
pattern <- "[\\d?.\\d]"
str_view_all(my_string, pattern)
```


### b.

Explain how you would use regular expression to count the number of words in x. Include the actual commands you would need to accomplish this task.

```{r}
str_replace_all(my_string, pattern = "[\\d?.\\d]", "") %>% 
  str_extract_all(pattern = "\\w+") %>%  
  unlist() %>%  
  length()
```


### c. 

Carefully explain what does the following code chunk will do.

```{r}
pattern <- "[\\d+]"
str_replace_all(my_string, pattern, "X")
```


### d.

Suppose you want to place a periods at the end of the `my_string`. The replacement attempt below contains one mistake. Write down the correct command that can place a periods at the end of `my_string`.

```{r}
pattern <- "$"
str_replace_all(my_string, pattern, ".")
```




