---
title: "Homework 7"
output: html_document
---

## Name: Put your name here

## I worked with: 

**Click the "Knit" button in RStudio to knit this file to a pdf.**

-----------------------------------------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, 
                      warning = FALSE, message = FALSE)
# add packages here 
library(tidyverse)
library(shiny)
library(nasaweather)
library(maps)
library(mapproj)
library(ggthemes)
theme_set(theme_stata(base_size = 10))  # for nice looking plots
```





## Problem 1:  Random guessing

Suppose you have $n$ observations and you would like to construct a classifier to predict positive outcomes. We know that $n_N$ are actually negative cases and $n_P$ are actually positive cases.

Suppose we are a lazy data scientist and don't actually build a model with predictors to classify our cases. Instead, we use a fair coin to classify negative and positive cases! This "coin classifier" will result in an even split between *expected* positive and negative predictions: 

$$
\hat{n}_N = \hat{n}_P = \dfrac{n}{2}
$$
Using the same logic, half of the actual failures will be successes and failures:
$$
TN = FP = \dfrac{n_N}{2}
$$
and half of the actual successes will be successes and failures:
$$
FN = TP = \dfrac{n_P}{2}
$$


confusion matrix          | truth negative | truth positive | total
------- | ---- | ------- | -----
predict negative    |    $TN=\dfrac{n_N}{2}$   |    $FN= \dfrac{n_P}{2}$   | $\hat{n}_N =\dfrac{n}{2}$
predict positive |   $FP=\dfrac{n_N}{2}$ |    $TP= \dfrac{n_P}{2}$   | $\hat{n}_P =\dfrac{n}{2}$
total           | $n_N$|  $n_P$  | $n$


### a.

*Answer:* Your answer here

```{r}
# Your code here
```



### b.

*Answer:* Your answer here

```{r}
# Your code here
```


### c. 

*Answer:* Your answer here

```{r}
# Your code here
```

### d. 

*Answer:* Your answer here

```{r}
# Your code here
```



------------------------------------------------------------------------------------------------



## Problem 3: Spam using k-nn

This example looks at a data set of about 4600 emails that are classified as spam or not spam, along with over 50 variables measuring different characteristic of the email. Details about these variables are found on the  [Spambase example ](http://archive.ics.uci.edu/ml/datasets/Spambase) on the machine learning data archive. The dataset linked to below is a slightly cleaned up version of this data. The only extra column in the data is `rgroup` which is a randomly assigned grouping variable (groups 0 through 99) which we will eliminate from the data.

Read the data in using the commands below to create a response `class` variable that contains the factor levels `spam` and `nonspam` with `spam` the first level. 

```{r}
# tsv = tab separated values!
spam <- read_delim("http://math.carleton.edu/kstclair/data/spamD.txt", 
        delim="\t")

# some clean up
spam <- spam %>% 
  mutate(class = fct_recode(
    spam, 
    spam = "spam" , 
    nonspam = "non-spam"), # rename levels because caret doesn't like "non-spam" 
    class = fct_relevel(class, "spam") # make "spam" the first level (our "positive")
    ) %>%  
  select(-rgroup, -spam) # don't need random group variable and  spam variable
levels(spam$class)  # verify "spam" is level 1
```

### a.

*Answer:* Your answer here

```{r}
# Your code here
```


### b.

```{r}
set.seed(757302859)  # set a seed
```

*Answer:* Your answer here

```{r}
# Your code here
```


### c. 

Make a recipe for fitting k nearest-neighbor algorithm to the training data by inputting the formula and the preprocessing steps.

*Answer:* Your answer here

```{r}
# Your code here
```


### d. 

*Answer:* Your answer here

```{r}
# Your code here
```

### e.

*Answer:* Your answer here

```{r}
# Your code here
``` 



### f.

*Answer:* Your answer here

```{r}
# Your code here
```


### g. 

*Answer:* Your answer here

```{r}
# Your code here
```


### h.

Use the `tidymodels` package to do 10-fold cross validation as follows:

- use the 80% training data split from part b.
- tune your knn spam classifier based on accuracy 
- consider neighborhood sizes ranging from size 1 to 31

Use the `results` to get the training set cross-validated estimates of the accuracy, precision, sensitivity and specificity of your final ("best") classifier.

And use the following seed before running your `train` command:

```{r}
set.seed(30498492)
```


*Answer:* Your answer here

```{r}
# Your code here
```

### i.

*Answer:* Your answer here

```{r}
# Your code here
```




-----------------------------------------------------------------------------------

## Problem 3: Logistic Regression


In this problem, you will create a classification model that can predict the presence of heart disease in individuals based on physical attributes and test results. By accurately predicting heart disease, the model can potentially help in reducing the need for more invasive diagnostic procedures.

```{r}
heart_data <- read_csv("https://raw.githubusercontent.com/deepbas/statdatasets/main/heart.csv")
```


### a.

```{r}
library(tidymodels)
set.seed(123322324)

# your code here

```



### b.


```{r}
# your code here

```


### c.


```{r}
# your code here

```


### d.

Explore the impact of varying classification thresholds on the prediction of heart disease. Plot the ROC curve, identifying the optimal threshold by evaluating sensitivity and specificity across a range of values. Discuss the clinical implications of selecting this threshold, particularly in terms of balancing false positives and false negatives.


```{r}
# your code here

```


