---
title: "Individual homework 7"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)

# List of required packages
library(tidyverse)
library(shiny)
library(dplyr)
library(babynames)
library(ggplot2)
```

## Please push your assignment to GitHub by 10:00 pm (Central) Tuesday, May 21. 

You are currently in the GitHub repository (repo) for `hw7-username`. The assignment prompt is shown below (i.e. in `README.Rmd`). You can view this online in your homework 7 GitHub repository as a Markdown file(`README.md`) or a pdf.

Please **use `hw7.Rmd` to complete this assignment**. Be sure to **knit your file to `pdf` before your final push to GitHub**.


## Homework process

For help on the homework process, review

-   [Assignments in Stat 220](https://stat220-spring24.netlify.app/assignments) for
    content/formatting questions.

-   [GitHub Guide for Students in Stat 220](https://stat220-spring24.netlify.app/github_tutorial)
    for Git and Github instructions.

When you are done with your homework, **don’t forget to push your changes to ALL files back to GitHub!** This means you should commit and push all related files, not just your final PDF. Additionally, ensure you post the link to your **GitHub repository to Gradescope for the final submission and grading**. This step is crucial as it allows for a comprehensive review of both your code and the rendered output


------------------------------------------------

## Assignment prompt

## Problem 1:  Random guessing

Suppose you have $n$ observations and you would like to construct a classifier to predict positive outcomes. We know that $n_N$ are actually negative cases and $n_P$ are actually positive cases.

Suppose we are a lazy data scientist and don't actually build a model with predictors to classify our cases. Instead, we use a fair coin to classify negative and positive cases! This "coin classifier" will result in an even split between *expected* positive and negative predictions: 

$$
\hat{n}_N = \hat{n}_P = \dfrac{n}{2}
$$
Using the same logic, half of the actual failures will be successes and failures:
$$
TN = FP = \dfrac{n_N}{2}
$$
and half of the actual successes will be successes and failures:
$$
FN = TP = \dfrac{n_P}{2}
$$


confusion matrix          | truth negative | truth positive | total
------- | ---- | ------- | -----
predict negative    |    $TN=\dfrac{n_N}{2}$   |    $FN= \dfrac{n_P}{2}$   | $\hat{n}_N =\dfrac{n}{2}$
predict positive |   $FP=\dfrac{n_N}{2}$ |    $TP= \dfrac{n_P}{2}$   | $\hat{n}_P =\dfrac{n}{2}$
total           | $n_N$|  $n_P$  | $n$


### a.

Use the "coin classifier" confusion matrix above to compute accuracy, precision, sensitivity, and specificity for this fair coin prediction method.

### b.

Consider another classification method that doesn't actually build a model with predictors to classify our cases. Instead, we now predict all cases to be positive! (e.g. a biased coin that lands "positive" with probability 1)

Construct the confusion matrix for this classifier and compute accuracy, precision, sensitivity, and specificity. 

### c. 

Under what conditions in part (b) will accuracy of this classifier be highest? Under what conditions in part (b) will accuracy of this classifier be lowest? 

### d. 

Describe the trade-off in part (b) in specificity vs sensitivity. What would these values be if instead of classifying every case as positive, you classified all cases as negative? 


## Problem 2: Spam using k-nn

This example looks at a data set of about 4600 emails that are classified as spam or not spam, along with over 50 variables measuring different characteristic of the email. Details about these variables are found on the  [Spambase example ](http://archive.ics.uci.edu/ml/datasets/Spambase) on the machine learning data archive. The dataset linked to below is a slightly cleaned up version of this data. The only extra column in the data is `rgroup` which is a randomly assigned grouping variable (groups 0 through 99) which we will eliminate from the data.

Read the data in using the commands below to create a response `class` variable that contains the factor levels `spam` and `nonspam` with `spam` the first level. 

```{r}
# tsv = tab separated values!
spam <- read_delim("https://raw.githubusercontent.com/deepbas/statdatasets/main/spam.txt", 
        delim="\t")

# some clean up
spam <- spam %>% 
  mutate(class = fct_recode(
    spam, 
    spam = "spam" , 
    nonspam = "non-spam"), # rename levels because caret doesn't like "non-spam" 
    class = fct_relevel(class, "spam") # make "spam" the first level (our "positive")
    ) %>%  
  select(-rgroup, -spam) # don't need random group variable and  spam variable
levels(spam$class)  # verify "spam" is level 1
```

### a.

What proportion of cases are spam in this data set? The variable `class` is the true classification of an email. 



### b.

We want to  fit a k-nn classifier for `spam` using the 57 quantitative predictors (columns 1-57) from the spam data to an 80%/20% training and test set split. Create the training and test sets to do this. Please use the following seed.

```{r}
set.seed(757302859)  # set a seed
```


### c. 

Make a recipe for fitting k nearest-neighbor algorithm to the training data by inputting the formula and the preprocessing steps.


### d. 

Specify the nearest neighbor model to fit a classification model using 10 neighbors.


### e.

Define the workflow object feeding in the recipe and model specification. Then, fit the model to the training data. 



### f.

Evaluate the model on the test dataset and predict `class` for the test data set. 



### g. 

Compute the accuracy, sensitivity, specificity and precision for predicting test set responses. Which rate(s) would you like to be high as a user of this spam filter?


### h.

Use the `tidymodels` package to do 10-fold cross validation as follows:

- use the 80% training data split from part b.
- tune your knn spam classifier based on accuracy 
- consider neighborhood sizes ranging from size 1 to 31

Use the `results` to get the training set cross-validated estimates of the accuracy, precision, sensitivity and specificity of your final ("best") classifier.

And use the following seed before running your `train` command:

```{r}
set.seed(30498492)
```


### i.

Use the `results` object from your CV fit from part (i) to plot the accuracy, precision, sensitivity and specificity for the values of $k$ that you tuned over. Comment on which neighborhood sizes are "best" in terms of these four metrics (your answer may depend on the metric).




## Problem 3: Logistic Regression


In this problem, you will create a classification model that can predict the presence of heart disease in individuals based on physical attributes and test results. By accurately predicting heart disease, the model can potentially help in reducing the need for more invasive diagnostic procedures.

```{r}
heart_data <- read_csv("https://raw.githubusercontent.com/deepbas/statdatasets/main/heart.csv")
```


### a.

Load and prepare the dataset. Perform necessary preprocessing steps such as handling missing values, normalizing data, and encoding categorical variables. Split the dataset into training and testing sets to ensure the model can be accurately evaluated. The variable `target` is the outcome with 0 for no presence of heart disease and 1 for presence of heart disease. Please use the given seed.

```{r}
library(tidymodels)
set.seed(123322324)
```



### b.

Fit a logistic regression model using the `tidymodels` framework in R. Use the `target` variable 'presence of heart disease' and include relevant predictors based on the dataset.



### c.

Plot the confusion matrix. Evaluate your model's performance by calculating metrics such as accuracy, sensitivity, specificity, and the Area Under the Curve (AUC). These metrics will provide insights into how well the model can distinguish between patients with and without heart disease.




### d.

Explore the impact of varying classification thresholds on the prediction of heart disease. Plot the ROC curve, identifying the optimal threshold by evaluating sensitivity and specificity across a range of values. Discuss the clinical implications of selecting this threshold, particularly in terms of balancing false positives and false negatives.



--------------------------------------------------------------------------------------