---
title: "Midterm I"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment=NA, fig.align = 'center')
knitr::opts_chunk$set(echo = TRUE, size = "small", collapse=TRUE, comment=NA, warning = FALSE, message = FALSE)

library(gapminder)
library(dplyr)
library(ggplot2)
library(lubridate)
library(forcats)
library(ggthemes)
library(tidyr)
data("gapminder")
select <- dplyr::select
theme_set(theme_stata(base_size = 10))
```

<center>  
Name:        
</center> 

<br>

## Total Points: 100   

<br>
<br>


*`Gapminder` data*

Data includes health and income outcomes for 142 countries from 1952 to 2007 in increments of 5 years. The variables in the dataset are `country`, `continent`, `year`, `lifeExp`, `pop`, and `gdpPercap`. The descriptions for the variables are:

  - `country` : name of the country, factor with 142 levels
  - `continent`: name of the continent, factor with 5 levels
  - `year` : ranges from 1952 to 2007 in increments of 5 years (12 distinct years)
  - `lifeExp`: life expectancy at birth, in years
  - `pop` : population
  - `gdpPercap` : GDP per capita (US$\$$, inflation-adjusted)


```{r}
glimpse(gapminder)
```

The distinct continents in the data are as follows:

```{r}
gapminder %>% pull(continent) %>% unique()
```


#


## Part 1: Data Wrangling *(10 points each)*

**What do the following code chunks do?** Provide a thorough and intuitive (3-5 sentences) description of the output from each of the following R chunks. The chunks produce a new data set. Please give the dimensions in addition to your description.  Write your descriptions in regular English, without using variable names.


### a. 


```{r, echo=FALSE, eval=FALSE}
gapminder %>%
  filter(year %in% c(1952, 2007)) %>%
  group_by(continent, year) %>%
  summarize(median_gdpPercap = median(gdpPercap),
            median_lifeExp = median(lifeExp),
            median_pop = median(pop)) 
```


```{r, eval=FALSE}
gapminder %>%
  filter(year %in% c(1952, 2007)) %>%
  group_by(continent, year) %>%
  summarize(median_gdpPercap = median(gdpPercap),
            median_lifeExp = median(lifeExp),
            median_pop = median(pop)) 
```


\vspace*{2.5in}


### b.

```{r, eval=FALSE}
gapminder %>%
  mutate(gdp_total = gdpPercap * pop) %>%
  group_by(continent, year) %>%
  summarize(gdp_continent = sum(gdp_total)) %>%
  pivot_wider(names_from = year, 
              values_from = gdp_continent, 
              names_prefix = "year_")
```

\vspace*{2.5in}


```{r, echo=FALSE, eval=FALSE}
gapminder %>%
  mutate(gdp_total = gdpPercap * pop) %>%
  group_by(continent, year) %>%
  summarize(gdp_continent = sum(gdp_total)) %>%
  pivot_wider(names_from = year, 
              values_from = gdp_continent, 
              names_prefix = "year_")
```


### c. 

```{r, eval=FALSE}
set.seed(143)
selected_countries <- gapminder %>%
  distinct(country, continent) %>%
  group_by(continent) %>%
  slice_sample(n = 1) %>%
  pull(country)

gapminder %>%
  filter(country %in% selected_countries) %>%
  filter(year %in% c(1952, 2007)) %>%
  group_by(country, year) %>%
  summarize(median_gdpPercap = median(gdpPercap),
            median_lifeExp = median(lifeExp),
            total_pop = sum(pop)) %>%
  pivot_longer(cols = -c(country, year), 
               names_to = "stat", 
               values_to = "value") 
```

\vspace*{5in}


```{r, echo=FALSE, eval=FALSE}
set.seed(143)
selected_countries <- gapminder %>%
  distinct(country, continent) %>%
  group_by(continent) %>%
  slice_sample(n = 1) %>%
  pull(country)

gapminder %>%
  filter(country %in% selected_countries) %>%
  filter(year %in% c(1952, 2007)) %>%
  group_by(country, year) %>%
  summarize(median_gdpPercap = median(gdpPercap),
            median_lifeExp = median(lifeExp),
            total_pop = sum(pop)) %>%
  pivot_longer(cols = -c(country, year), 
               names_to = "stat", 
               values_to = "value") 
```


### d. 


```{r, eval=FALSE}
set.seed(143)
gapminder %>%
  filter(continent == "Asia") %>%
  distinct(country) %>%
  slice_sample(n = 5) %>%
  inner_join(gapminder, by = "country") %>%
  filter(year %in% c(1952, 2007)) %>%
  group_by(country, year) %>%
  summarize(median_gdpPercap = median(gdpPercap),
            median_lifeExp = median(lifeExp),
            total_pop = sum(pop)) %>%
  pivot_longer(cols = -c(country, year), 
               names_to = "stat", 
               values_to = "value") %>%
  pivot_wider(names_from = year, 
              values_from = value, 
              names_prefix = "year_")
```

\vspace*{3in}


```{r, echo=FALSE, eval=FALSE}
set.seed(143)
gapminder %>%
  filter(continent == "Asia") %>%
  distinct(country) %>%
  slice_sample(n = 5) %>%
  inner_join(gapminder, by = "country") %>%
  filter(year %in% c(1952, 2007)) %>%
  group_by(country, year) %>%
  summarize(median_gdpPercap = median(gdpPercap),
            median_lifeExp = median(lifeExp),
            total_pop = sum(pop)) %>%
  pivot_longer(cols = -c(country, year), 
               names_to = "stat", 
               values_to = "value") %>%
  pivot_wider(names_from = year, 
              values_from = value, 
              names_prefix = "year_")
```


\newpage

## Part 2: Graphics *(15 points each)*


### a. The scatter plot below visualizes the relationship between GDP per capita and life expectancy of countries in the year 2007. What are 5 ways you could improve the aesthetics and readability of this plot by following best data visualization practices? Also write 5 code modifications on the space provided below:


```{r, out.width="80%"}
gapminder %>%
  filter(year == 2007) %>%
  ggplot(aes(x = gdpPercap, y = lifeExp)) + 
  geom_point()
```

*Answer:*


\vspace*{3in}



```{r, eval=FALSE}
gapminder %>%
  filter(year == 2007) %>%
  ggplot(aes(x = gdpPercap, y = lifeExp, color = continent)) + 
  geom_point(aes(size = pop)) 
  
  
  
  
  
  

  
  
  
  




  
    
```




```{r, echo=FALSE, eval=FALSE}
gapminder %>%
  filter(year == 2007) %>%
  ggplot() + 
  geom_point(aes(x = gdpPercap, y = lifeExp, color = continent), size = 2, alpha = 0.7) +
  scale_x_log10(labels = scales::dollar_format()) +
  scale_color_brewer(palette = "Set2", name = "Continent") +
  labs(title = "GDP per Capita vs Life Expectancy in 2007",
       subtitle = "Data from Gapminder",
       x = "GDP per Capita (log scale)",
       y = "Life Expectancy") +
  theme_minimal()
```


\newpage



### b. The partial code used to generate the plot below is given with placeholders for code snippet. Please provide the appropriate code snippet.

```{r, echo=FALSE}
gapminder %>%
  mutate(year_date = make_date(year)) %>%
  filter(year == 2007, continent != "Oceania") %>%
  group_by(continent) %>%
  mutate(avg_lifeExp = mean(lifeExp)) %>%
  ungroup() %>%
  mutate(continent_reordered = fct_reorder(continent, avg_lifeExp)) %>%
  ggplot(aes(x = gdpPercap, y = lifeExp, size = pop, color = continent)) +
  geom_point(alpha = 0.7) +
  labs(title = "Life Expectancy vs GDP per capita by Continent (2007)",
       x = "GDP per capita",
       y = "Life Expectancy") +
  scale_fill_brewer(palette = "Set1") +
  scale_x_log10(labels = scales::comma) +
  facet_wrap(~continent_reordered, ncol = 2) +
  theme(legend.position = "bottom") 
```



```{r, eval=FALSE}
gapminder %>%
  ##### FILL IN i. ##### %>%
  ##### FILL IN ii. ##### %>%
  group_by(continent) %>%
  mutate(avg_lifeExp = mean(lifeExp)) %>%
  ungroup() %>%
  ##### FILL IN iii. ##### %>%
  ##### FILL IN iv. ##### +
  geom_point(alpha = 0.7) +
  labs(title = "Life Expectancy vs GDP per capita by Continent (2007)",
       x = "GDP per capita",
       y = "Life Expectancy") +
  scale_fill_brewer(palette = "Set1") +
  scale_x_log10(labels = scales::comma) +
  ##### FILL IN v. ##### +
  theme(legend.position = "bottom") 
```


#### i. Create a new column called "year_date" that converts the "year" column into a date object 

\vspace*{1in}


#### ii. Filter the data to only include rows where the year is 2007 and exclude continent "Oceania"

\vspace*{1in}


#### iii. Reorder the levels of the "continent" factor based on the "avg_lifeExp" column and store the result in a new column called "continent_reordered."

\vspace*{1in}


#### iv. Create a ggplot2 scatter plot with "gdpPercap" on the x-axis, "lifeExp" on the y-axis, point sizes representing the "pop" column, and points color-coded by the "continent" column.


\vspace*{1in}


#### v. Create a faceted plot based on the "continent_reordered" column, with 2 columns of panes.

\vspace*{1in}



\newpage

## Part 3: Data Objects *(5 points each)*


```{r}
x <- 4:1
y <- c(TRUE, factor(c(NA, "b")), 1)
z <- list(z1 = x, z2 = y, z3 = c("Carleton", "college"), z4 = matrix(1:9, nrow = 3))
```


Consider the above objects to tell what each of the following code chunks evaluate to? Briefly explain your answer.

### (a)


```{r, eval=FALSE}
y[3]
```


\vspace*{0.8in}


### (b)



```{r, eval=FALSE}
z[["z3"]][1]   
```

\vspace*{0.8in}


### (c)


```{r, eval=FALSE}
z[x][[2]][[2]]
```

\vspace*{0.8in}


### (d)


```{r, eval=FALSE}
x - y
```

\vspace*{1in}


\newpage

### (e)


```{r, eval=FALSE}
unlist(z)
```

\vspace*{2in}


### (f)


```{r, eval=FALSE}
typeof(unlist(z))
```
