---
title: "Midterm I"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment=NA, fig.align = 'center')
knitr::opts_chunk$set(echo = TRUE, size = "small", collapse=TRUE, comment=NA, warning = FALSE, message = FALSE)

library(tidyverse)
library(gapminder)
library(ggthemes)
data("gapminder")
select <- dplyr::select
theme_set(theme_stata(base_size = 10))
```


Section I: Gapminder Data

Health and income outcomes for 142 countries from 1952 to 2007 in increments of 5 years. The variables in the dataset are `country`, `continent`, `year`, `lifeExp`, `pop`, and `gdpPercap`. The descriptions for the variables are:

  - `country` : name of the country, factor with 142 levels
  - `continent`: name of the continent, factor with 5 levels
  - `year` : ranges from 1952 to 2007 in increments of 5 years (12 distinct years)
  - `lifeExp`: life expectancy at birth, in years
  - `pop` : population
  - `gdpPercap` : GDP per capita (US$\$$, inflation-adjusted)


#

```{r}
glimpse(gapminder)
```


\newpage


## Part 1: What do the following codes do?

Provide a thorough and intuitive (2-3 sentences) description of the output from each of the following R chunks. The chunks either produce a new data set or a new plot; if it's a new data set, give the dimensions in addition to your description.  Write your descriptions in regular English, without using variable names.


### a.

```{r, eval=FALSE}
gapminder %>%
  filter(
    continent == "Africa",
    country == "Somalia",
     year %in% seq(1952, 2007, 5)) %>%
 slice_min(lifeExp, n=1)
```

\vspace*{2in}

### b.

```{r, eval=FALSE}
gapminder %>%
  filter(year == 2007) %>%
  group_by(continent) %>%
  summarise(totalPop = sum(pop)) %>%
  arrange(totalPop)
```

\vspace*{2in}


### c. 


```{r, eval=FALSE}
gapminder %>%
  group_by(country, continent) %>%
  summarize(meanPop = mean(pop)/1000000, meanLifeExp = mean(lifeExp)) %>%
  filter(continent == "Europe") %>%
  ggplot(aes(x = meanPop, y = meanLifeExp, color = country, label = country)) +
  geom_point() +
  scale_color_discrete(guide = "none") +
  ggrepel::geom_text_repel()
```

\vspace*{2in}


### d.

```{r, eval=FALSE}
gapminder %>%
  filter(year == 2007) %>%
  mutate(rank = min_rank(desc(lifeExp))) %>%
  filter(rank < 50) %>%
  arrange(rank) %>%
  slice_min(rank, n = 10) %>%
  select(-continent, -pop, -year, -gdpPercap)
```


\vspace*{2in}

\newpage

### e. 

```{r, eval=FALSE}
gapminder %>% 
  group_by(continent, year) %>%
  summarize(
      mean_le = mean(lifeExp),
      median_le = quantile(lifeExp, 0.50),
      min_le = min(lifeExp),
      max_le = max(lifeExp),
      se_le = sd(lifeExp)/sqrt(n()))%>%
 slice_max(mean_le, n=1) %>%
 ungroup()
```


\vspace*{2in}


## Part 2: Miscellaneous

### a. Use piping operation to combine the following set of codes into one chained line of code to produce the same output. Also, explain the output of the code.

```{r, eval=FALSE}
arrange(filter(gapminder, year == 1952), desc(gdpPercap))
```


```{r, echo=FALSE, eval=FALSE}
gapminder %>%
  filter(year == 1952) %>%
  arrange(desc(gdpPercap))
```


\vspace*{3in}


### b.  Use piping operation to combine the following set of codes into one chained line of code to produce the same output. Also, explain the output of the code.


```{r, eval=FALSE}
plot_data <- filter(gapminder, year == 1972 & continent %in% c("Africa", "Americas"))
ggplot(plot_data, aes(x = lifeExp, y = gdpPercap)) +
  geom_point() 
```


```{r, echo=FALSE, eval=FALSE}
gapminder %>% filter(year == 1972 & continent %in% c("Africa", "Americas")) %>%
  ggplot(aes(x = lifeExp, y = gdpPercap, color = continent)) +
  geom_point(aes(color = continent, size = pop)) +
  geom_smooth(se=FALSE)
```


\vspace*{2in}


### c. How would you improve the aesthetics of the plot given below? Write your code modifications.

```{r}
gapminder %>%
  filter(year == 2007) %>%
  ggplot(aes(x = gdpPercap, y = lifeExp), size = pop, color = continent) + 
  geom_point()
```


```{r, echo=FALSE, eval=FALSE}
gapminder %>%
  filter(year == 2007) %>%
  ggplot(aes(x = gdpPercap, y = lifeExp, size = pop, color = continent)) + 
  geom_point()
```


\vspace*{2in}


### d. Please take a look at the `# code 1` and the resulting plot. 

```{r}
# code 1
gapminder %>% 
  group_by(continent, year) %>%
  summarize(life_expectancy = lifeExp) %>%   # line 1
  ggplot(aes(x = continent, y = life_expectancy, color = continent)) +
  geom_boxplot() 
```

Will the plot look the same or different if we switch to `# code 2`? Explain.

```{r, eval=FALSE}
# code 2
gapminder %>% 
  group_by(continent, year) %>%
  summarize(life_expectancy = mean(lifeExp)) %>%
  ggplot(aes(x = continent, y = life_expectancy, color = continent)) +
  geom_boxplot() 
```


```{r, echo=FALSE, eval=FALSE}
gapminder %>% 
  group_by(continent, year) %>%
  summarize(life_expectancy = mean(lifeExp)) %>%
  ggplot(aes(x = continent, y=life_expectancy, color = continent)) +
  geom_boxplot() 
```


```{r, echo=FALSE, eval=FALSE}
gapminder %>% 
  group_by(continent, year) %>%
  summarize(life_expectancy = mean(lifeExp)) %>%
  ggplot(aes(x = continent, y=life_expectancy, color = continent)) +
  geom_boxplot() 
```





\vspace*{2.5in}

\newpage

Section II: R objects

## Part 3. Consider the following objects to answer the questions below. 

```{r}
x <- 1:4
y <- c(TRUE, as.factor(c(2,"3")), 4)
z <- list(z1 = x, z2 = y, z3 = c("cellar","door"))
```


#### (a)

Consider the objects x, y and z. Which are atomic vectors and which are lists?


\vspace*{2in}


#### (b)

What does the following command evaluate to? Briefly explain your answer.

```{r, eval = FALSE}
typeof(x)  # line 1
typeof(y) # line 2
```


\vspace*{2.5in}


#### (c)

What does the following command evaluate to? Briefly explain your answer.

```{r, eval=FALSE}
z[["z3"]][1]   
```


\vspace*{2in}


#### (d)

What does the following command evaluate to? Briefly explain your answer.

```{r, eval=FALSE}
z[3][[1]][2]    
```


\vspace*{2in}


#### (e)

What does the following command evaluate to? Briefly explain your answer.

```{r, eval=FALSE}
x + y
```


