---
title: "Class Activity 4"
author: "Your name here"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      size = "small", 
                      collapse = TRUE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      error = FALSE, # document will compile even in case of error
                      fig.width = 4,
                      fig.height= 3,
                      fig.align = 'center') 
# load the necessary libraries
library(tidyverse)
library(ggplot2)
library(datasauRus)
```



## Your turn 1

This worksheet will guide you through creating various plots using the `ggplot2` package in R. We will be using the `datasaurus`_dozen dataset from the `datasauRus` package for demonstration purposes. The dataset contains 13 different datasets, and we'll use them to create a variety of plots.

### Scatterplot

```{r, echo=FALSE}
# prepare the dataset
dino_data <- datasaurus_dozen %>%
   filter(dataset == "dino")
```

a. Run the following code.

```{r}
ggplot(data = dino_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

b. You _must_ remember to put the aesthetic mappings in the `aes()` function! What happens if you forget? 

*Answer:* 

If you forget to put the aesthetic mappings inside the aes() function, ggplot2 will not be able to map the variables to the aesthetics correctly, and you might encounter an error or unexpected behavior in your plot.

```{r}
# Add a layer and see what happens
ggplot(data = dino_data , x = x , y = y)
```


c. The aesthetic mappings can be specified in the geom layer if you prefer, instead of the main `ggplot()` call. Give it a try:

*Answer:* 

```{r}
# Rebuild the scatterplot with your aesthetic mapping in the geom layer
ggplot(data = dino_data) +
   geom_point(aes(x = x, y = y)) 
```


### Bar Plot

In this problem, we'll explore creating a bar plot using the `datasaurus`_dozen dataset.

a. Create a new data frame containing the count of observations in each dataset.

*Answer:* 

```{r}
dataset_counts <- datasaurus_dozen %>%
  group_by(dataset) %>%
  summarise(count = n())
```

b. Create a bar plot showing the number of observations in each dataset.

*Answer:* 

```{r}
ggplot(data = dataset_counts, aes(x = dataset, y = count, fill = dataset)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.key.size = unit(0.5, "cm")) 
```


```{r}
ggplot(data = dataset_counts, aes(x = dataset, fill = dataset)) +
  geom_bar(aes(y = after_stat(prop))) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.key.size = unit(0.5, "cm")) 
```


### Histogram

a. Create a histogram of the `x` variable for the `dino` dataset.

*Answer:* 

```{r}
ggplot(data = dino_data, aes(x = x)) +
  geom_histogram(binwidth = 7) 
```


b. Overlay a density curve on the histogram.

*Answer:* 

```{r}
ggplot(data = dino_data, aes(x = x)) +
  geom_histogram(aes(y = after_stat(density)), binwidth = 7, fill = "lightblue") +
  geom_density(color = "red")
```


### Boxplot

*Answer:* 

a. Create a boxplot of the x variable for each dataset in datasaurus_dozen.

```{r}
ggplot(data = datasaurus_dozen, aes(x = dataset, y = x)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


### Faceting

*Answer:* 

a. Create a scatterplot of `x` vs. `y` for each dataset in `datasaurus_dozen` using `facet_wrap()`.

```{r}
ggplot(data = datasaurus_dozen, aes(x = x, y = y)) +
  geom_point() +
  facet_wrap(~ dataset) +
  theme_minimal()
```


### Variable Transformation

a. The scatterplot of the `dino` dataset without any transformations is given below.

*Answer:* 

```{r}
ggplot(data = dino_data, aes(x = x, y = y)) +
  geom_point() +
  theme_minimal() -> p1
```


b. Now, apply the square root transformation to both the `x` and `y` axes using the `scale_x_sqrt()` and `scale_y_sqrt()` functions in the `dino` dataset.

*Answer:* 

```{r}
ggplot(data = dino_data, aes(x = x, y = y)) +
  geom_point() +
  scale_x_sqrt() +
  scale_y_sqrt() +
  theme_minimal() -> p2
```


c. Finally, use `grid.arrange()` function from `gridExtra` package to plot the above two plots side-by-side. Which plot do you prefer and why?

*Answer:* The second plot is more revealing of a dinosaur than the first plot. 

```{r}
library(gridExtra)
grid.arrange(p1, p2, nrow = 1)
```







