---
title: "Class Activity 12"
author: "Your name here"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      size = "small", 
                      collapse = TRUE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      error = TRUE)

# load the necessary libraries
library(stringr)
library(dplyr)
library(readr)
```


In this tutorial, we will learn about string manipulations using regular expressions and the `stringr` library in R. We will cover different examples and use cases to help you understand the concepts and functions related to string manipulation.


## Group Activity 1

```{r}
x <- "My SSN is 593-29-9502 and my age is 55"
y <- "My phone number is 612-643-1539"
z <- "My old SSN number is 39532 9423."
out <- str_flatten(c(x,y,z), collapse = ". ")
```

### a. What characters in `x` will `str_view_all(x, "-..-")` find?

*answer:*

The pattern: "-" then "something" then "something" then "-"

```{r}
str_view_all(x, "-..-")
```

### b. What pattern will `str_view_all(x, "-\\d{2}-")` find?

*answer:*


```{r}
str_view_all(x, "-\\d{2}-")  # "-" then 2 digits then "-"
```

### c. What pattern will `str_view_all(out, "\\d{2}\\.*")` find?

*answer:*


```{r}
str_view_all(out, "\\s\\d{2}\\.")  # 2 digits then "."
```


### d. Use `str_view_all` to determine the correct regex pattern to identify all SSN in `out`

*answer:*


We can get the SSN with the usual format (###-##-####) with a regex that has 3, 2, and 4 digits separated by a dash. 

```{r}
str_view_all(out,"([0-8]\\d{2})-(\\d{2})-(\\d{4})")
```

This misses the oddly formatted SSN in the third entry. Rather than use a dash, we can specify the divider as `[-\\s]?` which allows either 0 or 1 occurrences of either a dash or space divider:

```{r }
str_view_all(out,"([0-8]\\d{2})[-\\s]?(\\d{2})[-\\s]?(\\d{4})")
```


### e. Write a regular expression to extract dates in the format YYYY-MM-DD from a given text.

```{r}
date_pattern <- "\\d{4}-\\d{2}-\\d{2}"
text <- "The event will take place on 2023-07-20 and end on 2023-07-22."
str_extract_all(text, date_pattern)
```

*answer:*


### f. Write a regular expression to extract all words that start with a capital letter in a given text.

*answer:*


```{r}
capital_pattern <- "\\b[A-Z][a-zA-Z]*\\b"
text <- "Alice and Bob went to the Market to buy some Groceries."
str_extract_all(text, capital_pattern)
```



-----------------------------------------------------------------------------------

## Group Activity 2

### a. Let's deal with a number string that is longer than 9 digits.

```{r }
ssn <- "([0-8]\\d{2})[-\\s]?(\\d{2})[-\\s]?(\\d{4})"
test <- c("123-45-67890","1123 45 6789")
str_view_all(test, ssn)
```

This example captures a 9-digit string as an SSN, but these strings are longer than 9 digits and may not represent an SSN. One way to deal with this is to use the negative lookbehind `?<!` and negative lookahead `?!` operators to ensure that the identified 9-digit string does not have a leading 0 or does not contain more digits.

If we "look behind" from the start of the SSN, we should not see another digit:

```{r }
str_view_all(test, "(?<!\\d)([0-8]\\d{2})[-\\.\\s]?(\\d{2})[-\\.\\s]?(\\d{4})")
```

And if we "look ahead" from the end of the SSN, we should not see another digit:

```{r}
str_view_all(test, "(?<!\\d)([0-8]\\d{2})[-\\.\\s]?(\\d{2})[-\\.\\s]?(\\d{4})(?!\\d)")
```


For parts b and c, consider the following string. 

```{r}
string1 <- "100 dollars 100 pesos"
```

### b. Explain why the following matches the first 100 and not the second.

*answer:* 


```{r}
str_view(string1, "\\d+(?= dollars)")
```

### c. Explain why the following matches the second 100 and not the first.

*answer:* 


```{r}
str_view(string1, "\\d+(?!\\d| dollars)")
```


For parts d and e, please take a look at `string2`. 


```{r}
string2 <- "USD100 PESO100"
```

### d. Explain why the following matches the first 100 and not the second.

*answer:* 


```{r}
str_view(string2, "(?<=USD)\\d{3}")
```


### e. Explain why the following matches the second 100 and not the first.

*answer:* 


```{r}
str_view(string2, "(?<!USD)\\d{3}")
```


-------------------------------------------------------------------

##  Group Activity 3

```{r}
tweets<- read_csv("https://raw.githubusercontent.com/deepbas/statdatasets/main/TrumpTweetData.csv")
```

### a. What proportion of tweets (text) mention "America"?

```{r}
tweets %>% 
  summarize(prop = mean(str_detect(str_to_title(text), "America"))) 
```

### b. What proportion of these tweets include "great"?

```{r}
tweets %>% filter(str_detect(str_to_title(text), "America")) %>%
  summarize(prop = mean(str_detect(str_to_lower(text), "great")))   
```

### c. What proportion of the tweets mention `@`?

```{r}
tweets %>% mutate(ct = str_count(text, "@")) %>%
  select(text, ct) %>%
  summarize(prop = mean(ct>0))
```


### d. Remove the tweets having mentions `@`.

```{r}
Mentions <- c("@[^\\s]+")

tw_noMentions <- tweets %>% mutate(textNoMentions = str_replace_all(text, Mentions, ""))
tw_noMentions$text[38]
```

### e. What poportion of tweets originated from an iPhone?


```{r}
tweets %>% group_by(source) %>% summarize(count = n()) %>%
  mutate(prop = count / sum(count)) %>%  filter(source == "iPhone")
```




