---
title: "Class Activity 19"
author: "Your name here"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
output:
  html_document:
    df_print: paged
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      size = "small", 
                      collapse = TRUE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      error = TRUE) # change it to TRUE

# List of required packages
library(tidyverse)
library(shiny)
library(dplyr)
library(babynames)
```


## Question 1

We are utilizing the babynames dataset to explore and visualize how the popularity of specific names has evolved over time.


### (a) Create a user interface (UI) by using `textInput` for the baby name input, `selectInput` for gender selection, and adding an `actionButton` to initiate the search.

```{r}
ui <- fluidPage(
  titlePanel("Baby Names Trend"),
  sidebarLayout(
    sidebarPanel(
      textInput("name", "Enter a Baby Name:", "Emma"),
      selectInput("gender", "Select Gender:", choices = c("Male" = "M", "Female" = "F")),
      actionButton("goButton", "Show Trend")
    ),
    mainPanel(plotOutput("nameTrend"))
  )
)


```


### (b) Filter the `babynames` dataset based on user inputs for name and gender by utilizing `eventReactive` to respond to the action button press, and apply filter within this reactive context, using `isolate` to access the inputs without establishing reactive dependencies.


```{r}
server <- function(input, output) {
  # Use eventReactive to execute code in response to a specific action (button press)
  nameData <- eventReactive(input$goButton, {
    req(input$name) # Ensure the name input is not empty
    babynames %>% 
      filter(name == isolate(input$name), 
             sex == isolate(input$gender)) # Use isolate to prevent reactivity except on button press
  })
}


```


### (c) Create a line plot to visualize the trend of a selected baby name over the years by using `renderPlot`, showing the number of occurrences `(n)` of the name across different years `(year)`.

```{r}
server <- function(input, output) {
   # Use eventReactive to execute code in response to a specific action (button press)
  nameData <- eventReactive(input$goButton, {
    req(input$name) # Ensure the name input is not empty
    babynames %>% 
      filter(name == isolate(input$name), 
             sex == isolate(input$gender)) # Use isolate to prevent reactivity except on button press
  })
  
  output$nameTrend <- renderPlot({
   # req(nameData())  # Ensure data is available before plotting
    ggplot(nameData(), aes(x = year, y = n)) +
      geom_line() +  
      labs(title = paste("Trend for name", isolate(input$name)),  # Use isolate to get name input without triggering reactivity
           x = "Year", y = "Number of Babies") +
      theme_minimal()  
  })
}

```

Remember, without `isolate`, any change to the inputs `(input$name and input$gender)` would immediately trigger the reactive expressions (`nameData` and the plotting logic within `renderPlot`) to re-execute. This means the plot would update every time the user types a letter in the name input field or changes the gender, rather than waiting for the user to press the `goButton`.

### (d) Run the Shiny app by combining the UI and server components using `shinyApp.`

```{r}
shinyApp(ui, server, options = list(height = 800))
```



### (e). Deploying a Shiny App to `shinyapps.io`


Open an account on [shinyapps.io](https://www.shinyapps.io/admin/#/dashboard) and follow the steps to deploy one of the apps to [shinyapps.io](https://www.shinyapps.io/admin/#/dashboard).

#### Step 1: Create Your Shiny App

Create a file named app.R in a new folder. This folder will contain all the necessary files for your app. 

#### Step 2: Organize Your App Folder

Ensure your app.R file is saved in a dedicated folder for your app. This folder can also include any additional data files, data folder for data, or other resources your app needs.

#### Step 3: Deploy Your App to `shinyapps.io`

Before deploying, you'll need to set up an account on `shinyapps.io`. Use the rsconnect package to authenticate your R session with your `shinyapps.io` account. You will need the API key from your `shinyapps.io` account dashboard.

```r
library(rsconnect)
rsconnect::setAccountInfo(name='<account-name>',
                          token='<api-token>',
                          secret='<api-secret>')
```

Replace `<account-name>`, `<api-token>`, and `<api-secret>` with your actual account name and API key details. Navigate to your app folder in RStudio, or set the working directory to your app folder in R: 

```r
setwd("/path/to/your/app")
```

Deploy your app using the `deployApp` function:

```r
rsconnect::deployApp()
```

Follow the prompts in the R console. Once the deployment process completes, you will receive a URL to access your app hosted on `shinyapps.io`.

