---
title: "Class Activity 19"
author: "Your name here"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
output:
  html_document:
    df_print: paged
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      size = "small", 
                      collapse = TRUE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      error = TRUE) # change it to TRUE

# List of required packages
library(tidyverse)
library(shiny)
library(dplyr)
library(babynames)
```


\vspace*{1in}


Explore the magic firsthand! Simply paste the code provided in the slides into the field below and run it to witness the output in action.

```{r}
# Try the Shiny Code Yourself! 1
library(shiny)
ui <- fluidPage(
  titlePanel("Updating Plot Based on User Input"),
  sidebarLayout(
    sidebarPanel(
      numericInput("obs", "Number of observations:", 
                   value = 100, min = 1)
    ),
    mainPanel( plotOutput("distPlot"))
  )
)
server <- function(input, output) {
  output$distPlot <- renderPlot({
    hist(rnorm(input$obs))
  })
}
shinyApp(ui, server)
```


```{r}
# Try the Shiny Code Yourself! 2
ui <- fluidPage(
  mainPanel(
    actionButton("addButton", "Add 1"),
    textOutput("result")
  )
)
server <- function(input, output) {
  sum_so_far <- eventReactive(input$addButton, {
    if (is.null(input$addButton)) {
      0
    } else {
      input$addButton
    }
  })
  output$result <- renderText({
    paste("Sum so far:", sum_so_far())
  })
}
shinyApp(ui, server)
```

```{r}
# Try the Shiny Code Yourself! 3
ui <- fluidPage(
  mainPanel(
    numericInput("obs", "Number of observations", value = 100, min = 1),
    actionButton("updateButton", "Update Plot"),
    plotOutput("distPlot")
  )
)
server <- function(input, output) {
  data_to_plot <- eventReactive(input$updateButton, {
    rnorm(input$obs)
  })
  output$distPlot <- renderPlot({
    hist(data_to_plot())
  })
}
shinyApp(ui, server)
```

```{r}
# Try the Shiny Code Yourself! 4
ui <- fluidPage(
  titlePanel("Simple Counter"),
  numericInput("adjustValue", "Adjust by:", 1),
  actionButton("incrementButton", "Increment"),
  actionButton("decrementButton", "Decrement"),
  br(),
  textOutput("currentValue")
)
server <- function(input, output) {
  counter <- reactiveValues(value = 0)
  observeEvent(input$incrementButton, {
    counter$value <- counter$value + input$adjustValue
  })
  observeEvent(input$decrementButton, {
    counter$value <- counter$value - input$adjustValue
  })
  output$currentValue <- renderText({
    paste("Current Value:", counter$value)
  })
}
shinyApp(ui, server)
```

```{r}
# Try the Shiny Code Yourself! 5
ui <- fluidPage(
  mainPanel(
    actionButton("showAlert", "Show Alert"),
    textOutput("alertCount")
  )
)
server <- function(input, output) {
  alert_counter <- reactiveValues(count = 0)
  observeEvent(input$showAlert, {
    showModal(modalDialog(
      title = "Alert",
      "This is an alert message!"
    ))
    alert_counter$count <- alert_counter$count + 1
  })
  output$alertCount <- renderText({
    paste("Number of alerts shown:", alert_counter$count)
  })
}
shinyApp(ui, server)
```

```{r}
# Try the Shiny Code Yourself! 6
ui <- fluidPage(
  textInput("textInput", "Enter some text"),
  actionButton("submitButton", "Submit"),
  textOutput("outputText"),
  textOutput("counter")
)
server <- function(input, output) {
  counter <- reactiveValues(clicks = 0)
  observeEvent(input$submitButton, {
    counter$clicks <- counter$clicks + 1
    output$outputText <- renderText({
      # Using isolate() to prevent reactivity on textInput changes
      paste("You submitted:", isolate(input$textInput))
    })
    output$counter <- renderText({
      paste("Submit button clicked", counter$clicks, "times")
    })
  })
}
shinyApp(ui, server)
```



## Question 1

We are utilizing the `babynames` dataset to explore and visualize how the popularity of specific names has evolved over time in a Shiny app.


### (a) Create a user interface (UI) by using `textInput` for the baby name input, `selectInput` for gender selection, and adding an `actionButton` to initiate the search.

```{r}
# fill in the portions marked with # fill #
ui <- fluidPage(
  titlePanel("Baby Names Trend"),
  sidebarLayout(
    sidebarPanel(
      textInput("name", "Enter a Baby Name:", "Emma"),
      selectizeInput("gender", "Select Gender:", choices = c("Male" = "M", "Female" = "F")),
      actionButton("goButton", "Show Trend")
    ),
    mainPanel(plotOutput("nameTrend"))
  )
)
```


### (b) Filter the `babynames` dataset based on user inputs for name and gender by utilizing `eventReactive` to respond to the action button press, and apply filter within this reactive context, using `isolate` to access the inputs without establishing reactive dependencies.


```{r}
# fill in the portions marked with # fill #

server <- function(input, output) {
  nameData <- eventReactive(input$goButton, {
    req(input$name) # Ensure the name input is not empty
    babynames %>% 
      filter(name == input$name, sex == isolate(input$gender))
  })
}

shinyApp(ui, server, options = list(height = 800))


```


### (c) Create a line plot to visualize the trend of a selected baby name over the years by using `renderPlot`, showing the number of occurrences `(n)` of the name across different years `(year)` in addition to what you did in part b.

```{r}
server <- function(input, output) {
  nameData <- eventReactive(input$goButton, {
    req(input$name) # Ensure the name input is not empty
    babynames %>% 
      filter(name == isolate(input$name), sex == isolate(input$gender))
  })
  
  
  output$nameTrend <- renderPlot({
    req(nameData()) 
    ggplot(nameData(), aes(x = year, y = n)) +
      geom_line() +
      labs(title = paste("Trend for name", isolate(input$name)),
           x = "Year", y = "Number of Babies") +
      theme_minimal()
  })
}
```

Remember, without `isolate`, any change to the inputs `(input$name and input$gender)` would immediately trigger the reactive expressions (`nameData` and the plotting logic within `renderPlot`) to re-execute. This means the plot would update every time the user types a letter in the name input field or changes the gender, rather than waiting for the user to press the `goButton`.

### (d) Run the Shiny app by combining the UI and server components using `shinyApp.`

```{r}
shinyApp(ui, server, options = list(height = 800))
```


## Select multiple names


```{r}
ui <- fluidPage(
  titlePanel("Baby Names Trends for Multiple Names"),
  sidebarLayout(
    sidebarPanel(
      textAreaInput("names", "Enter Baby Names (separated by commas):", "Emma,Olivia,Noah"),
      selectizeInput("gender", "Select Gender:", choices = c("Male" = "M", "Female" = "F")),
      actionButton("goButton", "Show Trends")
    ),
    mainPanel(plotOutput("nameTrends"))
  )
)

```


```{r}
server <- function(input, output) {
  nameData <- eventReactive(input$goButton, {
    req(input$names) # Ensure the name input is not empty
    inputNames <- unlist(strsplit(isolate(input$names), ",")) %>% str_trim() 
    
    data <- lapply(inputNames, function(inputName) {
      babynames %>%
        filter(name == inputName, sex == isolate(input$gender))
    }) %>% bind_rows() # Combine data for all names
    
    return(data)
  })
  
  output$nameTrends <- renderPlot({
    req(nameData())
    ggplot(nameData(), aes(x = year, y = n, color = name)) +
      geom_line() +
      labs(title = "Trend for Names",
           x = "Year", y = "Number of Babies") +
      theme_minimal()
  })
}

```


```{r}
shinyApp(ui, server, options = list(height = 800))
```

