---
title: "Class Activity 3"
author: "Your name here"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
format:
  pdf: default
  html:
    df_print: paged
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
  error: true
---


```{r}
# load the necessary libraries
library(tidyverse)
library(ggplot2)
library(babynames)
```


```{r}
# some interesting data objects
x <- c(3,6,9,5,10)
x.mat <- cbind(x, 2*x)
x.df <- data.frame(x=x,double.x=x*2)
my.list <- list(myVec=x, myDf=x.df, myString=c("hi","bye"))
```

### Question 1: data types

- What data type is `x`? 

*Answer:*

```{r}
# code
typeof(x)
typeof(babynames$number)
```

- What data type is `c(x, x/2)`?

*Answer:*

```{r}
# code
typeof(c(x, babynames$year))
```

- What data type is `c(x,NA)`?  What data type is `c(x,"NA")`?


*Answer:*

```{r}
# code
typeof(c(x, NA))
typeof(c(x, "NA"))
```


### Question 2: Subsetting and coercion

- How can we reverse the order of entries in `x`?

*Answer:*

```{r}
# code
rev(x)
x[length(x):1]
```


- What does `which(x < 5)` equal?

*Answer:*

```{r}
# code
which(x<5)
```


- Extract the element of x that corresponds to the location in the preceding question.

*Answer:*

```{r}
# code
x[which(x<5)]
```



- What does `sum(c(TRUE,FALSE,TRUE,FALSE))` equal?

*Answer:*

```{r}
# code
sum(c(TRUE,FALSE,TRUE,FALSE))
```

- What does `sum(x[c(TRUE,FALSE,TRUE,FALSE)])` equal?

*Answer:*

```{r}
# code
sum(x[c(TRUE,FALSE,TRUE,FALSE, TRUE)])
```

- What does `sum(x < 5)` equal?

*Answer:*

```{r}
# code
sum(x < 5)
```

- What does `sum(x[x < 5])` equal?


*Answer:*

```{r}
# code
sum(x[x < 5])
```

- Why `dim(x.mat[1:2,1])` return `NULL` while `dim(x.mat[1:2,1:2])` returns a dimension?

*Answer:*

```{r}
# code
dim(x.mat[1:2,1])
dim(x.mat[1:2,1:2])
```


### Question 3: Lists

- Using `my.list`, show three ways to write one command that gives the 3rd entry of variable `x` in data frame `myDf`

*Answer:*

```{r}
# code
my.list[[1]][3]
my.list[["myVec"]][3]
my.list[1]$myVec[3]
my.list$myVec[3]
```


- What class of object does the command `my.list[3]` return?

*Answer:*

```{r}
# code
class(my.list[3])
```

- What class of object does the command `my.list[[3]]` return?

*Answer:*

```{r}
# code
class(my.list[[3]])
```

- What class of object does the command `unlist(my.list)` return? Why are all the entries `character`s?

*Answer:*

```{r}
# code
class(unlist(my.list))
```



