---
title: "Class Activity 17"
author: "Your name here"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
format:
  pdf: default
  html:
    df_print: paged
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
  error: true
---

```{r setup, include=FALSE}
# load the necessary libraries
library(tidyverse)
library(stringr)
library(purrr)
library(ggthemes)
library(rvest)
library(polite)
```


\vspace*{1in}


## Group Activity 1

1. Go to the [the numbers webpage](https://www.the-numbers.com/movie/budgets/all) and extract the table on the front page.

```{r}
session1 <- bow(url = "https://www.the-numbers.com/movie/budgets/all") %>% scrape() %>%
  

```


2. Find out the number of pages that contain the movie table, while looking for the changes in the url in the address bar. How does the url changes when you go to the next page?

*Answer:* 


3. Write a for loop to store all the data in multiple pages to a single data frame. Also, do the same using `purrr::map_df `.

```{r}
library(tidyverse)
library(rvest)

new_urls <- "https://www.the-numbers.com/movie/budgets/all/"

# Create an empty list
df1 <- list()

# Generate a vector of indices
index <- seq(1, 6301, 100)
```



## Group Activity 2

1. Go to [scrapethissite](https://www.scrapethissite.com/pages/forms/) and extract the table on the front page.

```{r}
session1 <- read_html("https://www.scrapethissite.com/pages/forms/") 
```


2. Find out the number of pages that contain the movie table, while looking for the changes in the url in the address bar. How does the url changes when you go to the next page?



3. Write a for loop to store all the data in multiple pages to a single data frame. Also, do the same using `purrr::map_df `.



