---
title: "Class Activity 17"
author: "Your name here"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      size = "small", 
                      collapse = TRUE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      error = TRUE)

# load the necessary libraries
library(tidyverse)
library(stringr)
library(purrr)
library(ggthemes)


# Set ggplot theme
theme_set(theme_stata(base_size = 8))
```


\vspace*{1in}


## Group Activity 1

1. Go to the [the numbers webpage](https://www.the-numbers.com/movie/budgets/all) and extract the table on the front page.

```{r}
session1 <- bow(url = "https://www.the-numbers.com/movie/budgets/all") %>% scrape() %>%
  html_elements(css = "table") %>%
  html_table()

table_base <- session1[[1]]
```


2. Find out the number of pages that contain the movie table, while looking for the changes in the url in the address bar. How does the url changes when you go to the next page?

*Answer:* The starting count of the movie get concatenated to the url in increments of 100.

3. Write a for loop to store all the data in multiple pages to a single data frame.

```{r}
new_urls <- "https://www.the-numbers.com/movie/budgets/all/"

# creating two empty data-frames

table_new <-data.frame()
df2 <- data.frame()


idx <-  seq(1, 6301, 100)

for (i in 1:length(idx)) {
  new_webpage <- read_html(str_glue(new_urls, {idx[i]}))
  table_new <- html_table(new_webpage)[[1]] %>% 
  tibble::as_tibble(.name_repair = "unique") 
  df2 <- rbind(df2, table_new)
}
```


```{r, eval=FALSE}
# Alternate using bow, takes very long
df3 <- data.frame()

urls <- map(idx, function(i) str_glue("https://www.the-numbers.com/movie/budgets/all/", {i}))
sessions <- lapply(1:length(urls), function(i) bow(urls[[i]]) %>% scrape() %>% html_elements(css = "table") %>% html_table())
data <- do.call(rbind, lapply(1:64, function(i) sessions[[i]][[1]]))
```


4. Display the data table using `knitr::kable(data, format = "html")`

```{r}
knitr::kable(head(df2), format = "html")
```


5. Store the data in a `.csv` file with `write_csv(data_object, "/path/name.csv")`

```{r}
readr::write_csv(df2, "my_data.csv")
```


## Group Activity 2

```{r}
mca <- data.frame(percent = c(23.5, 34.7, 19.8, 22.0, 27.6, 32.1, 20.2, 20.2, 26.0, 30.7, 22.1, 21.1), 
                  portfolio = factor(rep(c("Commodities", "Stocks", "Real State", "Other Assets"), 3), 
                                     levels=c("Commodities", "Stocks", "Real State", "Other Assets")), 
                  year = rep(c("2020","2021","2022"), each=4))
glimpse(mca)
```



```{r}
mybar <- ggplot(mca, aes(x=year, y=percent, fill=portfolio)) + 
  labs(title="Investment Portfolio") +
  geom_bar(stat = "identity", position = "dodge") +
 # theme(legend.position = "bottom") + 
  scale_fill_wsj()
ggplotly(mybar) 
```

